---
title: "8-2 meeting"
author: "lulu zhang"
date: "8/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sas7bdat)
library(tidyr)
library(tidyverse)
library(SASxport)
library(dplyr)
require(haven)
library(car)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
require(ggplot2)
library(janitor)
library(ggplot2)
require(sandwich)
require(msm)
library(forcats)
require(ggplot2)
require(gridExtra)
```

```{r load.data., include=FALSE}
nhes1 = read.csv("./nhesI_per.csv") %>% 
  janitor::clean_names()
nhes2 = read.csv("./nhesII_per.csv")%>% 
  janitor::clean_names()
nhes3 = read.csv("./nhesIII_per.csv")%>% 
  janitor::clean_names()
nhanes1 = read.csv("./nhanes1_per.csv")%>% 
  janitor::clean_names()  %>% 
  mutate(weight_all = coalesce(all_samp_weight_1_35, all_sampweight_1_65, detailed_weight_66_100, detailed_weight_1_35, detailed_weight_1_65))
n1_smkhbp = read.csv("./n1hbpsmk.csv") %>%
  janitor::clean_names() %>% 
  mutate(hbpx = coalesce(hypertension, hbp)) %>% 
  select(seqn, ever_smoker,  hbpx) %>% rename(hbp = hbpx)

nhanes1 = merge(nhanes1, n1_smkhbp, by = "seqn", all = TRUE)

nhanes2 = read.csv("./nhanes2_per.csv")%>% 
  janitor::clean_names() %>% 
  mutate(race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
         race = fct_explicit_na(race, na_level = "Other"),
         gender = dplyr::recode(gender, `0` = "Male",
                                `1` = "Female"))
n2hbp = read.csv("./n2hbp.csv") %>% janitor::clean_names() %>% 
mutate(hbp = dplyr::recode(hbp,
                           `No` = "0",
                           `Yes`= "1"),
       hbp = as.numeric(as.character(hbp)))

nhanes2 = merge(nhanes2, n2hbp, by = "seqn", all = TRUE)
nhanes3 = read.csv("./nhanes3_per.csv")%>% 
  janitor::clean_names()
newnhanes = read.csv("./new_per.csv")%>% 
  janitor::clean_names()
nhis = read.csv("./nhis_per.csv")%>% 
  janitor::clean_names() %>% 
  mutate(race = as.factor(race),
         gender = as.factor(gender),
         race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
        gender = dplyr::recode(gender, `0` = "Male",
                               `1` = "Female")) %>% 
  rename(hbp = hypertension)


```


## bmi
```{r bmi, message=FALSE, echo=FALSE}
hes2bmi = nhes2  %>%
  select(age_grp, period, race, gender, bmi, weight, psu, strata)
hes3bmi = nhes3 %>%
  select(age_grp, period, race, gender, bmi, weight, psu, strata)
hanes1bmi = nhanes1 %>%
  select(age_grp, period, race, gender, bmi, weight, psu, strata)
hanes2bmi = nhanes2 %>%
  select(age_grp, period, race, gender, bmi, weight, psu, strata)
hanes3bmi = nhanes3 %>%
  select(age_grp, period, race, gender, bmi, weight, psu, strata)
newhanesbmi = newnhanes %>%
  select(age_grp, period, race, gender, bmi, weight, psu, strata)
nhisbmi = nhis%>%
  select(age_grp, period, race, gender, bmicalc, weight, psu, strata) %>% 
  rename(bmi = bmicalc)

a = rbind(hes2bmi, hes3bmi, hanes1bmi)
b = rbind(hanes2bmi, hanes3bmi, newhanesbmi)
bmi_all = rbind(a, b, nhisbmi)%>% filter(age_grp %in% c(0:9)) %>% 
mutate(
          age_grp = as.factor(age_grp),
          age_grp = dplyr::recode(age_grp,
                                  `0` = "0-4",
                                  `1` = "5-9",
                                  `2` = "10-14",
                                  `3` = "15-19",
                                  `4` = "20-24",
                                  `5` = "25-29",
                                  `6` = "30-34",
                                  `7` = "35-39",
                                  `8` = "40-44",
                                  `9` = "45-49"))

bmi_all = bmi_all  %>% 
  filter(period %in% c(0:11),
         between(bmi, 10,50), 
         race %in% c("White", "Black", "Other"),
         gender %in% c("Male", "Female")) %>% 
   group_by(age_grp, period, race, gender) %>% 
   mutate(weighted_bmi = weighted.mean(bmi, weight))

## plot
bmi_all %>% 
ggplot(aes(x=period, weighted_bmi, color=age_grp))+
  geom_line()+
  facet_wrap(~gender+race, scales = "free_y")+
  labs(title = "weighted means of bmi from 1963-2018 among 0-49 year olds",
       ylab = "average bmi") +  scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



```
Peak for female age 20-24, other race period 2 (1970-1974) bc 7 obs with one at 37.3 bmi and one at 31.7 with a greater weight, other obs around 17-20

## hypertension

```{r hbp, echo=FALSE, message=FALSE}

## subset and merge datasets with hbp as a variable
hes1hbp = nhes1 %>%
  select(age_grp, period, race, gender, hbp, weight)
hanes1hbp = nhanes1 %>%
  select(age_grp, period, race, gender, hbp, weight, psu, strata)
hanes2hbp = nhanes2%>%
  select(age_grp, period, race, gender, hbp, weight, psu, strata)
hanes3hbp = nhanes3 %>%
  select(age_grp, period, race, gender, hbp, weight, psu, strata)
newhaneshbp = newnhanes %>%
  select(age_grp, period, race, gender, hbp, weight, psu, strata) 
nhishbp = nhis%>%
  select(age_grp, period, race, gender, hbp, weight, psu, strata)

a = rbind(hanes2hbp, hanes3hbp, newhaneshbp, nhishbp)
hbp = full_join(hes1hbp, a) 

# export to sas to get weighted means for responses
hbpmeans = read_xlsx("./bmi_means.xlsx", sheet = 9) %>% 
  filter(age_grp %in% c(0:9))
  
hbpmeans = hbpmeans%>%   mutate(race = as.factor(race),
         gender = as.factor(gender),
         age_grp = as.factor(age_grp),
         age_grp = dplyr::recode(age_grp,
                                 `0` = "0-4",
                                  `1` = "5-9",
                                  `2` = "10-14",
                                  `3` = "15-19",
                                  `4` = "20-24",
                                  `5` = "25-29",
                                  `6` = "30-34",
                                  `7` = "35-39",
                                  `8` = "40-44",
                                  `9` = "45-49"),
         VarLevel = as.numeric(as.factor(VarLevel))) ## 1 = no, 2 == yes


## plot 'no hbp'
hbpmeans %>% 
  filter(VarLevel == 1,
         Mean < 1)%>% 
  ggplot( aes(x = period, y = Mean))+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted proportion with no hypertension among 0-49 year olds by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")+scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## plot 'yes' hbp
hbpmeans %>% 
  filter(VarLevel == 2,
         Mean < 1)%>% 
  ggplot( aes(x = period, y = Mean))+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted proportion with hypertension among 0-49 year olds by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")+  scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

## thyroid disease

```{r thyroid,  echo=FALSE, message=FALSE}
hanes1thy = nhanes1%>%
  select(age_grp, period, race, gender, thyroid_disease, weight_all, psu, strata) %>% 
  rename(ever_had_thyroid_d = thyroid_disease,
         weight = weight_all)
hanes2thy = nhanes2%>%
  select(age_grp, period, race, gender, ever_had_thyroid_d, weight, psu, strata)
hanes3thy = nhanes3%>%
  select(age_grp, period, race, gender, ever_had_thyroid_d, weight, psu, strata)
newhanesthy = newnhanes%>%
  select(age_grp, period, race, gender, ever_had_thyroid_d, weight, psu, strata)
nhisthy = nhis%>%
  select(age_grp, period, race, gender, ever_had_thyroid_d, weight, psu, strata)

thy = rbind(hanes1thy, hanes2thy, hanes3thy, newhanesthy, nhisthy)


thymeans = read_xlsx("./bmi_means.xlsx", sheet = 10)%>% mutate(age_grp = as.numeric(age_grp)) %>% 
  filter(age_grp %in% c(0:9))
  
thymeans = thymeans%>%   mutate(race = as.factor(race),
         gender = as.factor(gender),
         age_grp = as.factor(age_grp),
         age_grp = dplyr::recode(age_grp,
                                 `0` = "0-4",
                                  `1` = "5-9",
                                  `2` = "10-14",
                                  `3` = "15-19",
                                  `4` = "20-24",
                                  `5` = "25-29",
                                  `6` = "30-34",
                                  `7` = "35-39",
                                  `8` = "40-44",
                                  `9` = "45-49")) ## 1 = no, 2 == yes

thymeans %>% 
  filter(VarLevel == 0,
         Mean < 1)%>% 
  ggplot( aes(x = period, y = Mean))+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted proportion with no thyroid disease among 0-49 year olds by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")+scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## plot 'yes' thyroid
thymeans %>% 
  filter(VarLevel == 1,
         Mean < 1)%>% 
  ggplot( aes(x = period, y = Mean))+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted proportion with thyroid disease among 0-49 year olds by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")+  scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```
Other Male and female has a lot of 1.0000s for age groups 15-29 for period 2, 
15-34, 40-49 for period 3, 20-44 for period 5, very low sample sizes, 0.0000 and 1.0000s were not plotted 

## ghp 

```{r ghp.means, echo=FALSE, message=FALSE}
ghp = read_xlsx("./bmi_means.xlsx", sheet = 6) 
ghp = ghp %>% 
  mutate(period = as.factor(period),
         AGE_GRP = as.factor(AGE_GRP),
         race = as.factor(race),
         gender = as.factor(gender),
         period = dplyr::recode(period,
                                `1988-1994` = "1",
                                `1999-2002` = "2",
                                `2003-2006` = "3",
                                `2007-2010` = "4",
                                `2011-2014` = "5",
                                `2015-2018` = "6"),
         period = as.numeric(period))

ghp %>% ggplot(aes(x=period, y=Mean, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~gender+race, scales = "free_y")+
  labs(title = "weighted means of ghp % from 1988-2018 among 0-49 year olds",
       ylab = "average % ghp") +  scale_x_continuous(breaks=c(1,2,3,4,5,6),
  labels=c(`1` = "1988-1994",
           `2` = "1999-2002",`3` = "2003-2006", `4` = "2007-2010", `5` = "2011-2014", `6` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## diabetes
```{r diabetes, echo = FALSE, message=FALSE}
nhes1diabetes = nhes1 %>% 
  select(weight, race, gender, diabetes, age_grp, period)
nhes2diabetes = nhes2 %>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)
nhes3diabetes = nhes3 %>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)
hanes1diabetes = nhanes1%>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)
hanes2diabetes = nhanes2%>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)
hanes3diabetes = nhanes3%>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)
newdiabetes = newnhanes%>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)
nhisdiabetes = nhis%>% 
  select(weight, race, gender, diabetes, age_grp, period, strata, psu)

d = rbind(nhes2diabetes, nhes3diabetes, hanes1diabetes, hanes2diabetes, hanes3diabetes, newdiabetes, nhisdiabetes)
diabetes = full_join(nhes1diabetes, d)

diabetesmeans = read_xlsx("./bmi_means.xlsx", sheet = 8)
## plots 
diabetesmeans = diabetesmeans%>%  filter(age_grp %in% c(0:9)) %>% 
  mutate(race = as.factor(race),
         gender = as.factor(gender),
         age_grp = as.factor(age_grp),
         age_grp = dplyr::recode(age_grp,
                                 `0` = "0-4",
                                  `1` = "5-9",
                                  `2` = "10-14",
                                  `3` = "15-19",
                                  `4` = "20-24",
                                  `5` = "25-29",
                                  `6` = "30-34",
                                  `7` = "35-39",
                                  `8` = "40-44",
                                  `9` = "45-49"),
        VarLevel = as.numeric(as.factor(VarLevel))) 
 

diabetesmeans%>% 
  filter(VarLevel == 1,
         gender %in% c("Male", "Female"),
         race %in% c("Black", "White", "Other"))%>% 
  ggplot( aes(x = period, y = Mean))+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted proportion that do not have diabetes among 0-49 year olds by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")+  scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

diabetesmeans%>% 
  filter(VarLevel == 2)%>% 
  ggplot( aes(x = period, y = Mean))+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted proportion with diabetes among 0-49 year olds by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")+  scale_x_continuous(breaks=c(0,1,2,3,4,5,6,7,8,9,10,11),
  labels=c(`0` = "1960-1964",`1` = "1965-1969",
           `2` = "1970-1974",`3` = "1975-1979",`4` = "1980-1984", `5` = "1985-1989", `6`= "1990-1994",
           `7` = "1995-1999", `8` = "2000-2004", `9` = "2005-2009", `10` = "2010-2014", `11` = "2015-2018"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```
Did not plot weighted means that were 0 or 1 due to small N's
NHESII:6-11 year olds
NHESIII:12-17 year olds

## smoker

```{r smoker, echo=FALSE, message=FALSE}
nhes3smoker = nhes3 %>% 
  select(race, gender, age_grp, num_cigs_smoked_daily, strata, weight, period, psu) %>% 
  rename(current_smoker = num_cigs_smoked_daily)
hanes1smoker = nhanes1 %>% 
  select(race, gender, age_grp, ever_smoker, strata, weight, period, psu) %>% 
  rename(current_smoker = ever_smoker) %>% 
  mutate(current_smoker = dplyr::recode(current_smoker,
                                        `2` = " "))
hanes2smoker = nhanes2%>% 
  select(race, gender, age_grp, current_smoker, strata, weight, period, psu)
hanes3smoker = nhanes3%>% 
  select(race, gender, age_grp, current_smoker, strata, weight, period, psu)
newsmoker = newnhanes%>% 
  select(race, gender, age_grp, current_smoker, strata, weight, period, psu)
nhissmoker = nhis%>% 
  select(race, gender, age_grp, current_smoker, strata, weight, period, psu)

smoker = rbind(nhes3smoker, hanes1smoker, hanes2smoker, hanes3smoker, newsmoker, nhissmoker)

```

## household smokers
```{r}
nhanes3 %>% 
   select(race, gender, age_grp, home_smokers, strata, weight, period, psu)
newnhanes %>% 
   select(race, gender, age_grp, home_smokers, strata, weight, period, psu)
  

```

## alcohol

## diet variables in quintiles

## univariate analyses
```{r}

```

