---
title: "9-23 by race and gender"
author: "lulu zhang"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sas7bdat)
library(tidyr)
library(tidyverse)
library(SASxport)
library(dplyr)
require(haven)
library(car)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
require(ggplot2)
library(janitor)
library(ggplot2)
require(sandwich)
require(msm)
library(forcats)
require(ggplot2)
require(gridExtra)
library(rlang)
library(dplyr)
library(BBmisc)
library(caret)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)
library(readxl)
```

```{r load.data, include=FALSE}
seerrf=read.csv("./seerrf.csv") %>% 
  mutate(period = as.factor(period))
```

### glm by race & gender 
 
```{r subset.race, include=FALSE}
wm =seerrf %>% filter(race == "White",
                   gender == "Male")

bm =seerrf %>% filter(race == "Black",
                   gender == "Male")

om =seerrf %>% filter(race == "Other",
                   gender == "Male")

wf =seerrf %>% filter(race == "White",
                   gender == "Female")

bf =seerrf %>% filter(race == "Black",
                   gender == "Female")

of =seerrf %>% filter(race == "Other",
                   gender == "Female")
```


### bmi

# white female
```{r bmi.glm, echo=FALSE}

wfbmi <- glm(tot_count ~ age_grp + period, data=wf, family=poisson(), offset = lnpop)
bmi1wf <- glm(tot_count ~ age_grp + period+ bmimean1, data=wf, family=poisson(), offset = lnpop)
bmi2wf <- glm(tot_count ~ age_grp + period+ bmimean2, data=wf, family=poisson(), offset = lnpop)
bmi12wf <- glm(tot_count ~ age_grp + period+ bmimean1+bmimean2, data=wf, family=poisson(), offset = lnpop)
summary(wfbmi)
summary(bmi1wf)
summary(bmi2wf)
summary(bmi12wf)

##plot betas from the model fit
plot_model(wfbmi, transform = NULL)
plot_model(bmi1wf, transform = NULL)
plot_model(bmi2wf, transform = NULL)
plot_model(bmi12wf, transform = NULL)
```
white females, negalitve, some p-values = 0.05

# white male
```{r wmbmi, echo=FALSE}
wmbmi <- glm(tot_count ~ age_grp + period, data=wm, family=poisson(), offset = lnpop)
bmi1wm <- glm(tot_count ~ age_grp + period+ bmimean1, data=wm, family=poisson(), offset = lnpop)
bmi2wm <- glm(tot_count ~ age_grp + period+ bmimean2, data=wm, family=poisson(), offset = lnpop)
bmi12wm <- glm(tot_count ~ age_grp + period+ bmimean1+bmimean2, data=wm, family=poisson(), offset = lnpop)
summary(wmbmi)
summary(bmi1wm)
summary(bmi2wm)
summary(bmi12wm)

##plot betas from the model fit
plot_model(wmbmi, transform = NULL)
plot_model(bmi1wm, transform = NULL)
plot_model(bmi2wm, transform = NULL)
plot_model(bmi12wm, transform = NULL)
```
white males, negative but not significant


# black female
```{r bfbmi, echo=FALSE}
bfbmi <- glm(tot_count ~ age_grp + period, data=bf, family=poisson(), offset = lnpop)
bmi1bf <- glm(tot_count ~ age_grp + period+ bmimean1, data=bf, family=poisson(), offset = lnpop)
bmi2bf <- glm(tot_count ~ age_grp + period+ bmimean2, data=bf, family=poisson(), offset = lnpop)
bmi12bf <- glm(tot_count ~ age_grp + period+ bmimean1+bmimean2, data=bf, family=poisson(), offset = lnpop)
summary(bfbmi)
summary(bmi1bf)
summary(bmi2bf)
summary(bmi12bf)

##plot betas from the model fit
plot_model(bfbmi, transform = NULL)
plot_model(bmi1bf, transform = NULL)
plot_model(bmi2bf, transform = NULL)
plot_model(bmi12bf, transform = NULL)
```



# black male
```{r bmbmi, echo=FALSE}
bmbmi <- glm(tot_count ~ age_grp + period, data=bm, family=poisson(), offset = lnpop)
bmi1bm <- glm(tot_count ~ age_grp + period+ bmimean1, data=bm, family=poisson(), offset = lnpop)
bmi2bm <- glm(tot_count ~ age_grp + period+ bmimean2, data=bm, family=poisson(), offset = lnpop)
bmi12bm <- glm(tot_count ~ age_grp + period+ bmimean1+bmimean2, data=bm, family=poisson(), offset = lnpop)
summary(bmbmi)
summary(bmi1bm)
summary(bmi2bm)
summary(bmi12bm)

##plot betas from the model fit
plot_model(bmbmi, transform = NULL)
plot_model(bmi1bm, transform = NULL)
plot_model(bmi2bm, transform = NULL)
plot_model(bmi12bm, transform = NULL)
```
black males betas negative but not significant



# other female
```{r ofbmi, echo=FALSE}
ofbmi <- glm(tot_count ~ age_grp + period, data=of, family=poisson(), offset = lnpop)
bmi1of <- glm(tot_count ~ age_grp + period+ bmimean1, data=of, family=poisson(), offset = lnpop)
bmi2of <- glm(tot_count ~ age_grp + period+ bmimean2, data=of, family=poisson(), offset = lnpop)
bmi12of <- glm(tot_count ~ age_grp + period+ bmimean1+bmimean2, data=of, family=poisson(), offset = lnpop)
summary(ofbmi)
summary(bmi1of)
summary(bmi2of)
summary(bmi12of)

##plot betas from the model fit
plot_model(ofbmi, transform = NULL)
plot_model(bmi1of, transform = NULL)
plot_model(bmi2of, transform = NULL)
plot_model(bmi12of, transform = NULL)
```
other females: betas mostly negative, none are significant

# other male
```{r ombmi, echo=FALSE}
ombmi <- glm(tot_count ~ age_grp + period, data=om, family=poisson(), offset = lnpop)
bmi1om <- glm(tot_count ~ age_grp + period+ bmimean1, data=om, family=poisson(), offset = lnpop)
bmi2om <- glm(tot_count ~ age_grp + period+ bmimean2, data=om, family=poisson(), offset = lnpop)
bmi12om <- glm(tot_count ~ age_grp + period+ bmimean1+bmimean2, data=om, family=poisson(), offset = lnpop)
summary(ombmi)
summary(bmi1om)
summary(bmi2om)
summary(bmi12om)

##plot betas from the model fit
plot_model(ombmi, transform = NULL)
plot_model(bmi1om, transform = NULL)
plot_model(bmi2om, transform = NULL)
plot_model(bmi12om, transform = NULL)
```
other males: obese is postive with -value = 0.05, overweight is negative and not significant in combined model, neither is significant in separate models


## ghp

# white female
```{r wfghp,  echo=FALSE}
wfghp <- glm(tot_count ~ age_grp + period, data=wf, family=poisson(), offset = lnpop)
ghpwf <- glm(tot_count ~ age_grp + period+ Zvarghp, data=wf, family=poisson(), offset = lnpop)
summary(wfghp)
summary(ghpwf)

##plot betas from the model fit
plot_model(wfghp, transform = NULL)
plot_model(ghpwf, transform = NULL)
```
white female: `-0.08217` pvalue = 0.05

# white male
```{r wmghp, echo=FALSE}
wmghp <- glm(tot_count ~ age_grp + period, data=wm, family=poisson(), offset = lnpop)
ghpwm <- glm(tot_count ~ age_grp + period+ Zvarghp, data=wm, family=poisson(), offset = lnpop)
summary(wmghp)
summary(ghpwm)

##plot betas from the model fit
plot_model(wmghp, transform = NULL)
plot_model(ghpwm, transform = NULL)
```
white male `-0.02722` not significant 

#black female
```{r bfghp, echo=FALSE}
wfghp <- glm(tot_count ~ age_grp + period, data=wf, family=poisson(), offset = lnpop)
ghpwf <- glm(tot_count ~ age_grp + period+ Zvarghp, data=wf, family=poisson(), offset = lnpop)
summary(wfghp)
summary(ghpwf)

##plot betas from the model fit
plot_model(wfghp, transform = NULL)
plot_model(ghpwf, transform = NULL)
```
black female: `-0.08217` pvalue = 0.05

#black male
```{r bmghp, echo=FALSE}
wmghp <- glm(tot_count ~ age_grp + period, data=wm, family=poisson(), offset = lnpop)
ghpwm <- glm(tot_count ~ age_grp + period+ Zvarghp, data=wm, family=poisson(), offset = lnpop)
summary(wmghp)
summary(ghpwm)

##plot betas from the model fit
plot_model(wmghp, transform = NULL)
plot_model(ghpwm, transform = NULL)
```
black male: `-0.02722` not significant

# other female
```{r ofghp, echo=FALSE}
wfghp <- glm(tot_count ~ age_grp + period, data=wf, family=poisson(), offset = lnpop)
ghpwf <- glm(tot_count ~ age_grp + period+ Zvarghp, data=wf, family=poisson(), offset = lnpop)
summary(wfghp)
summary(ghpwf)

##plot betas from the model fit
plot_model(wfghp, transform = NULL)
plot_model(ghpwf, transform = NULL)
```
other female: `-0.08217` pvalue = 0.05


#other male
```{r omghp, echo=FALSE}
wmghp <- glm(tot_count ~ age_grp + period, data=wm, family=poisson(), offset = lnpop)
ghpwm <- glm(tot_count ~ age_grp + period+ Zvarghp, data=wm, family=poisson(), offset = lnpop)
summary(wmghp)
summary(ghpwm)

##plot betas from the model fit
plot_model(wmghp, transform = NULL)
plot_model(ghpwm, transform = NULL)
```
other male: `-0.02722` not significant
