---
title: "9-23 meeting-by gender only"
author: "lulu zhang"
date: "9/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sas7bdat)
library(tidyr)
library(tidyverse)
library(SASxport)
library(dplyr)
require(haven)
library(car)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
require(ggplot2)
library(janitor)
library(ggplot2)
require(sandwich)
require(msm)
library(forcats)
require(ggplot2)
require(gridExtra)
library(rlang)
library(dplyr)
library(BBmisc)
library(caret)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)
library(readxl)
```


```{r load.data, echo=FALSE, message=FALSE, include=FALSE}
seer = read.csv("./groupedseer.csv") %>% 
  janitor::clean_names()
seerrf=read.csv("./seerrf.csv") %>% 
  mutate(period = as.factor(period))
seergender = read.csv("./seergender.csv")

allrf = read.csv("./seerrf.csv")%>% 
  janitor::clean_names() 

malerf = read.csv("./seermrf.csv")%>% 
  janitor::clean_names()

femrf = read.csv("./seerfrf.csv")%>% 
  janitor::clean_names()

```


```{r seer.inc, echo=FALSE, message=FALSE}
seer %>%
  filter(gender == "Female") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = race)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
    title = "CRC inc rates from 1975-2014 among Females",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
  labels=c(`1` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))

seer %>%mutate(period = as.numeric(period)) %>%
  filter(gender == "Male") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = race)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
    title = "CRC inc rates from 1975-2014 among males ",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(3,4,5,6,7,8,9,10,11),
  labels=c(`3` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))
```

## seer incidence trends among males and females by age groups
```{r, echo=FALSE, message=FALSE}
seerf = seergender %>% filter(gender == "Female") %>% 
  mutate(period = as.factor(period))
seerm = seergender %>% filter(gender == "Male")%>% 
  mutate(period = as.factor(period))

seerf%>%mutate(period = as.numeric(period),
               age_grp = as.factor(age_grp),
               age_grp = dplyr::recode(age_grp,
                                       `6` = "25-29",
                                       `7` = "30-34",
                                       `8` = "35-39",
                                       `9` = "40-44",
                                       `10` = "45-49"))  %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = age_grp)) +
   labs(
    title = "CRC inc rates from 1975-2014 among females ",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
  labels=c(`1` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))

seerm%>%mutate(period = as.numeric(period),
               age_grp = as.factor(age_grp),
               age_grp = dplyr::recode(age_grp,
                                       `6` = "25-29",
                                       `7` = "30-34",
                                       `8` = "35-39",
                                       `9` = "40-44",
                                       `10` = "45-49"))  %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = age_grp)) +
   labs(
    title = "CRC inc rates from 1975-2014 among males ",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
  labels=c(`1` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))
```

normalize all continuous risk factors for glm!
then plot beta estimates acorss periods, with seer incidence rates overlayed
then another panel with age groups as x-axis 

**plot_model, transform = NULL keeps betas, doesnt give IRR

```{r bmiplot, echo=FALSE, message=FALSE}
## this is just betas from bmi from excel
femalebmi = read_xlsx("./univariate models.xlsx", sheet = 8) %>%
  janitor::clean_names() %>% 
  mutate(beta_estimate = as.numeric(beta_estimate),
         period = as.factor(period))

 age_grp <- femalebmi$age_grp
ggplot(femalebmi, aes(x = period, y = beta_estimate)) + 
  geom_point()+
  labs(title = "beta estimates for period and bmi levels-overweight and obese
       for females")+
geom_hline(yintercept=age_grp, col = "red")

malebmi = read_xlsx("./univariate models.xlsx", sheet = 9) %>%
  janitor::clean_names() %>% 
  mutate(beta_estimate = as.numeric(beta_estimate),
         period = as.factor(period))
 age_grp <- malebmi$age_grp
ggplot(malebmi, aes(x = period, y = beta_estimate)) + 
  geom_point()+
  geom_line(aes(x = period, y = beta_estimate ))+
  labs(title = "beta estimates for period and bmi levels-overweight and obese
       for males") +
geom_hline(yintercept=age_grp, col = "red")

```

## try to overlay seer incidence over beta estimates
This is a plot of the model of beta coefficients for ghp female and male respectively
first and third plot are without rf, second and fourth are with RF
```{r ghp, echo=FALSE, message=FALSE}
ghpctsmean = read.csv("./ghpctsmean.csv")
ghpseer = merge(ghpctsmean, seer, by= c("period", "age_grp", "race", "gender"), all = TRUE) %>% 
  mutate(period = as.factor(period))
ghpseer$period <- relevel(ghpseer$period, ref = "9")
fghp = ghpseer %>% 
  filter(gender == "Female")
mghp = ghpseer %>% 
  filter(gender == "Male")

#female
femghp <- glm(tot_count ~ age_grp + period, data=fghp, family=poisson(), offset = lnpop)
ghpf <- glm(tot_count ~ age_grp + period+ Zvarghp, data=fghp, family=poisson(), offset = lnpop)
summary(femghp)
summary(ghpf)

##plot betas from the model fit
plot_model(femghp, transform = NULL)
plot_model(ghpf, transform = NULL)

#male
maleghp <- glm(tot_count ~ age_grp + period, data=mghp, family=poisson(), offset = lnpop)
ghpm <- glm(tot_count ~ age_grp + period+ Zvarghp, data=mghp, family=poisson(), offset = lnpop)
summary(maleghp)
summary(ghpm)

##plot betas from the model fit
plot_model(maleghp, transform = NULL)
plot_model(ghpm, transform = NULL)



```
females: `0.03392` p-value =0.1
males: `0.01976` p-value not sig


# plot bmi using plot_model function
```{r bmi.plot, echo=FALSE, message=FALSE}
female = seerrf %>% 
  filter(gender == "Female")
male = seerrf %>% 
  filter(gender == "Male")

female$period <- relevel(female$period, ref = "9")

male$period <- relevel(male$period, ref = "9")
#female
seerbase <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
seerbmi1 <- glm(tot_count ~ age_grp + period+bmimean1, data=female, family=poisson(), offset = lnpop)
seerbmi2 <- glm(tot_count ~ age_grp + period+bmimean2, data=female, family=poisson(), offset = lnpop)
seerbmi23 <- glm(tot_count ~ age_grp + period+bmimean1 +bmimean2, data=female, family=poisson(), offset = lnpop)
summary(seerbase)
summary(seerbmi1)
summary(seerbmi2)
summary(seerbmi23)
plot_model(seerbase)
plot_model(seerbmi1)
plot_model(seerbmi2)
plot_model(seerbmi23)


### males
seerbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
seerbmi1m <- glm(tot_count ~ age_grp + period+bmimean1, data=male, family=poisson(), offset = lnpop)
seerbmi2m <- glm(tot_count ~ age_grp + period+bmimean2, data=male, family=poisson(), offset = lnpop)
seerbmi23m <- glm(tot_count ~ age_grp + period+bmimean1 +bmimean2, data=male, family=poisson(), offset = lnpop)
summary(seerbasem)
summary(seerbmi1m)
summary(seerbmi2m)
summary(seerbmi23m)
##baseline model
plot_model(seerbasem)
##just add overweight
plot_model(seerbmi1m)
## just add obese
plot_model(seerbmi2m)
## add both overweight and obese
plot_model(seerbmi23m)
```
females: overweight: `0.2331`  pvalue = 0.05
         obese: `1.1592`  pvalue < 0.001 
         both:`0.40173` p-value not significant `0.88458` pvalue = 0.001  
males:   overweight: `0.06799` but not significant
         obese: `0.64594` pvalue  = 0.01 
         both:betas = `-0.59642` and `1.657976` respectively, both with pvalues < 0.000


##trying to get two y axis
tried to plot beta estimates and incidence on 2 y-xis for female bmi but having trouble plotting the age groups by groups 
```{r, echo=FALSE, message=FALSE}
library(plotly)
femalebmi = femalebmi %>% 
  mutate(period = as.factor(period)) 
seerf

x = full_join(femalebmi, seerf) %>% 
  mutate(age_grp = as.factor(age_grp))
obese <- x$obese
 plot_ly(x)%>%
    add_trace(x=~period,y=~rate,name = 'incidence rate',type = 'scatter',mode = 'lines+markers', connectgaps = TRUE, group_by = age_grp) %>%
    add_trace(x=~period,y=~beta_estimate,name = 'beta estimates',type = 'scatter',mode = 'lines+markers',connectgaps = TRUE,yaxis = "y2") %>%
    layout(title = 'bmi beta estimates with incidence rates',
       xaxis = list(title = "incidence rates"),
       yaxis2 = list(side = 'right', overlaying = "y", title = 'beta estimates', showgrid = FALSE, zeroline = FALSE))
 

```

## diabetes plot model
```{r diabetes, echo=FALSE, message=FALSE}
#####female
diabetesbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
diabetesyesf <- glm(tot_count ~ age_grp + period+diabetesmean, data=female, family=poisson(), offset = lnpop)
summary(diabetesbasef)
summary(diabetesyesf)
plot_model(diabetesbasef)
plot_model(diabetesyesf)
####male
diabetesbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
diabetesyesm <- glm(tot_count ~ age_grp + period+diabetesmean, data=male, family=poisson(), offset = lnpop)
summary(diabetesbasem)
summary(diabetesyesm)
plot_model(diabetesbasem)
plot_model(diabetesyesm)
```
females, diabetes beta = `2.117081` pvalue < 0
males, beta = `1.852686` pvalue < 0.01


## hbp 
```{r hbp}
#####female
hbpbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
hbpyesf <- glm(tot_count ~ age_grp + period+hbpmean, data=female, family=poisson(), offset = lnpop)
summary(hbpbasef)
summary(hbpyesf)
plot_model(hbpbasef)
plot_model(hbpyesf)
####male
hbpbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
hbpyesm <- glm(tot_count ~ age_grp + period+hbpmean, data=male, family=poisson(), offset = lnpop)
summary(hbpbasem)
summary(hbpyesm)
plot_model(hbpbasem)
plot_model(hbpyesm)
```

females: beta `1.526870` pvalue < 0
males: beta `1.23044` pvalue < 0



## alc per week tertiles
```{r alc.trt, echo=FALSE, message=FALSE}
#####female
alcbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
alc1f <- glm(tot_count ~ age_grp + period+alcmean1t, data=female, family=poisson(), offset = lnpop)
alc2f <- glm(tot_count ~ age_grp + period+alcmean2t, data=female, family=poisson(), offset = lnpop)
alc23f <- glm(tot_count ~ age_grp + period+alcmean1t+alcmean2t, data=female, family=poisson(), offset = lnpop)
summary(alcbasef)
summary(alc1f)
summary(alc2f)
summary(alc23f)
plot_model(alcbasef)
plot_model(alc1f)
plot_model(alc2f)
plot_model(alc23f)
####male
alcbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
alc1m <- glm(tot_count ~ age_grp + period+alcmean1t, data=male, family=poisson(), offset = lnpop)
alc2m <- glm(tot_count ~ age_grp + period+alcmean2t, data=male, family=poisson(), offset = lnpop)
alc23m <- glm(tot_count ~ age_grp + period+alcmean1t+alcmean2t, data=male, family=poisson(), offset = lnpop)
summary(alcbasem)
summary(alc1m)
summary(alc2m)
summary(alc23m)
plot_model(alcbasem)
plot_model(alc1m)
plot_model(alc2m)
plot_model(alc23m)
```
female: tert 1: `3.3922` p-value not sig
        tert 2: `-8.73245` p-value = 0.05
        both:  `8.73245` p-value = 0.05 & NA for tert 2
male:   tert 1: `0.27470`  p-value not sig
        tert 2: `0.881117` p-value not sig
        both: `4.61802` & `5.44251` p-values not sig
error note: *Model matrix is rank deficient. Parameters period3 were not estimable.Model matrix is rank deficient. Parameters age_grp were not estimable.Model matrix is rank deficient. Parameters age_grp, alcmean2t were not estimable*
-apparently it means not enough data points?


## thyroid-go back to this 
```{r thyroid, echo=FALSE, message=FALSE}
#####female
thybasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
thyyesf <- glm(tot_count ~ age_grp + period+thymean1, data=female, family=poisson(), offset = lnpop)
summary(thybasef)
summary(thyyesf)
plot_model(thybasef)
plot_model(thyyesf)
####male
thybasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
thyyesm <- glm(tot_count ~ age_grp + period+thymean1, data=male, family=poisson(), offset = lnpop)
summary(thybasem)
summary(thyyesm)
plot_model(thybasem)
plot_model(thyyesm)
```
female: `-1.03625` p-value < 0
male: `-2.06222`  p-value < 0.01


## meat per week-continuous and normalized
```{r cts.meat, echo=FALSE, message=FALSE}
ctsmeatbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
ctsmeatf <- glm(tot_count ~ age_grp + period+Zvarmeatmean, data=female, family=poisson(), offset = lnpop)
summary(ctsmeatbasef)
summary(ctsmeatf)
plot_model(ctsmeatbasef)
plot_model(ctsmeatf)
####male
ctsmeatbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
ctsmeatm <- glm(tot_count ~ age_grp + period+Zvarmeatmean, data=male, family=poisson(), offset = lnpop)
summary(ctsmeatbasem)
summary(ctsmeatm)
plot_model(ctsmeatbasem)
plot_model(ctsmeatm)
```
female: `0.020368` p-value not sig
male: `-0.18063` p-value < 0


## current smoker-yes
Female, then male 

```{r smk, echo=FALSE, message=TRUE}
smkbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
smkf <- glm(tot_count ~ age_grp + period+smkmean1, data=female, family=poisson(), offset = lnpop)
summary(smkbasef)
summary(smkf)
plot_model(smkbasef)
plot_model(smkf)
####male
smkbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
smkm <- glm(tot_count ~ age_grp + period+smkmean1, data=male, family=poisson(), offset = lnpop)
summary(smkbasem)
summary(smkm)
plot_model(smkbasem)
plot_model(smkm)
```
female: `0.44640` p-value not sig
male: `-0.02905` p-value not sig


## fiber, tertiles
female then male
```{r fiber, echo=FALSE, message=FALSE}
#####female
fibbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
fib1f <- glm(tot_count ~ age_grp + period+fib1mean, data=female, family=poisson(), offset = lnpop)
fib2f <- glm(tot_count ~ age_grp + period+fib2mean, data=female, family=poisson(), offset = lnpop)
fib3f <- glm(tot_count ~ age_grp + period+fib3mean, data=female, family=poisson(), offset = lnpop)
fib4f <- glm(tot_count ~ age_grp + period+fib4mean, data=female, family=poisson(), offset = lnpop)
fib14f <- glm(tot_count ~ age_grp + period+fib1mean+fib2mean+fib3mean+fib4mean, data=female, family=poisson(), offset = lnpop)
summary(fibbasef)
summary(fib1f)
summary(fib2f)
summary(fib3f)
summary(fib4f)
summary(fib14f)
plot_model(fibbasef)
plot_model(fib1f)
plot_model(fib2f)
plot_model(fib3f)
plot_model(fib4f)
plot_model(fib14f)
####male
fibbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
fib1m <- glm(tot_count ~ age_grp + period+fib1mean, data=female, family=poisson(), offset = lnpop)
fib2m <- glm(tot_count ~ age_grp + period+fib2mean, data=female, family=poisson(), offset = lnpop)
fib3m <- glm(tot_count ~ age_grp + period+fib3mean, data=female, family=poisson(), offset = lnpop)
fib4m <- glm(tot_count ~ age_grp + period+fib4mean, data=female, family=poisson(), offset = lnpop)
fib14m <- glm(tot_count ~ age_grp + period+fib1mean+fib2mean+fib3mean+fib4mean, data=female, family=poisson(), offset = lnpop)
summary(fibbasem)
summary(fib1m)
summary(fib2m)
summary(fib3m)
summary(fib4m)
summary(fib14m)
plot_model(fibbasem)
plot_model(fib1m)
plot_model(fib2m)
plot_model(fib3m)
plot_model(fib4m)
plot_model(fib14m)
```
female: quintile 1: `0.74232` p-value < 0.05
        quintile 2: `-1.02493` p-value < 0
        quintile 3: `-1.44639` p-value < 0
        quintile 4: `-0.27829` p-value not sig
        all:`-0.25427` p-value not sig; ` -0.80995` p-value =0.05; `-1.38366` p-value < 0; `-0.21427` p-value not sig
male:   quintile 1: `0.74232`  p-value = 0.05
        quintile 2: `-1.02493` p-value < 0
        quintile 3: `-1.44639` p-value < 0
        quintile 4: `-0.27829` p-value not sig
        all: `-0.25427` p-value not sig; `-0.80995` p-value = 0.05; `-1.38366` p-value < 0; `-0.21427` p-value not sig

## protein
```{r protein, echo = FALSE, message=FALSE}
#####female
protbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
prot1f <- glm(tot_count ~ age_grp + period+prot1mean, data=female, family=poisson(), offset = lnpop)
prot2f <- glm(tot_count ~ age_grp + period+prot2mean, data=female, family=poisson(), offset = lnpop)
prot12f <- glm(tot_count ~ age_grp + period+prot1mean+prot2mean, data=female, family=poisson(), offset = lnpop)
summary(protbasef)
summary(prot1f)
summary(prot2f)
summary(prot12f)
plot_model(protbasef)
plot_model(prot1f)
plot_model(prot2f)
plot_model(prot12f)
####male
protbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
prot1m <- glm(tot_count ~ age_grp + period+prot1mean, data=male, family=poisson(), offset = lnpop)
prot2m <- glm(tot_count ~ age_grp + period+prot2mean, data=male, family=poisson(), offset = lnpop)
prot12m <- glm(tot_count ~ age_grp + period+prot1mean+prot2mean, data=male, family=poisson(), offset = lnpop)
summary(protbasem)
summary(prot1m)
summary(prot2m)
summary(prot12m)
plot_model(protbasem)
plot_model(prot1m)
plot_model(prot2m)
plot_model(prot12m)
```
female: tert 1: `0.740271` p-value < 0
        tert 2: `0.072968` p-value not sig
        both: `0.41500` p-value = 0.1 & `-0.21583` p-value = 0.05
mal:    tert 1: `1.257292` p-value < 0
        tert 2: `-0.306681` p-value < 0 
        both: `1.48662` p-value < 0 & `0.23015` p-value = 0.05

## fat- tertile
```{r fat}
#####female
fatbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
fat1f <- glm(tot_count ~ age_grp + period+fat1mean, data=female, family=poisson(), offset = lnpop)
fat2f <- glm(tot_count ~ age_grp + period+fat2mean, data=female, family=poisson(), offset = lnpop)
fat12f <- glm(tot_count ~ age_grp + period+fat1mean+fat2mean, data=female, family=poisson(), offset = lnpop)
summary(fatbasef)
summary(fat1f)
summary(fat2f)
summary(fat12f)
plot_model(fatbasef)
plot_model(fat1f)
plot_model(fat2f)
plot_model(fat12f)
####male
fatbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
fat1m <- glm(tot_count ~ age_grp + period+fat1mean, data=male, family=poisson(), offset = lnpop)
fat2m <- glm(tot_count ~ age_grp + period+fat2mean, data=male, family=poisson(), offset = lnpop)
fat12m <- glm(tot_count ~ age_grp + period+fat1mean+fat2mean, data=male, family=poisson(), offset = lnpop)
summary(fatbasem)
summary(fat1m)
summary(fat2m)
summary(fat12m)
plot_model(fatbasem)
plot_model(fat1m)
plot_model(fat2m)
plot_model(fat12m)
```
female: tert 1: `-0.352030` p-value not sig
        tert 2: `0.624416` p-value < 0
        both: `-0.405388` & `-0.053022` p-value both not sig
male:   tert 1: `0.276805` p-value not sig
        tert 2: `-0.563022` p-value < 0
        both: `-1.352290` & `-1.143851` both p-value < 0  



## saturated fat- tertiles
try continuous saturated fat as well
```{r}
#####female
satfatbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
satfat1f <- glm(tot_count ~ age_grp + period+satfat1mean, data=female, family=poisson(), offset = lnpop)
satfat2f <- glm(tot_count ~ age_grp + period+satfat2mean, data=female, family=poisson(), offset = lnpop)
satfat12f <- glm(tot_count ~ age_grp + period+satfat1mean+satfat2mean, data=female, family=poisson(), offset = lnpop)
summary(satfatbasef)
summary(satfat1f)
summary(satfat2f)
summary(satfat12f)
plot_model(satfatbasef)
plot_model(satfat1f)
plot_model(satfat2f)
plot_model(satfat12f)
####male
satfatbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
satfat1m <- glm(tot_count ~ age_grp + period+satfat1mean, data=male, family=poisson(), offset = lnpop)
satfat2m <- glm(tot_count ~ age_grp + period+satfat2mean, data=male, family=poisson(), offset = lnpop)
satfat12m <- glm(tot_count ~ age_grp + period+satfat1mean+satfat2mean, data=male, family=poisson(), offset = lnpop)
summary(satfatbasem)
summary(satfat1m)
summary(satfat2m)
summary(satfat12m)
plot_model(satfatbasem)
plot_model(satfat1m)
plot_model(satfat2m)
plot_model(satfat12m)
```
female: tert 2: `0.265451` p-value = 0.05
        tert 3: `-0.158180` p-value not sig
        both: `0.32037` & `0.06747` both p-value not sig
male:   tert 2: `1.180400` p-value < 0
        tert 3: `-0.709995` p-value < 0
        both: `0.803931` p-value = 0.1 & `-0.303046` p-value not sig


## fat-continuous&normalized
```{r, echo=FALSE, message=FALSE}
ctsfatbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
ctsfatf <- glm(tot_count ~ age_grp + period+Zvarfat, data=female, family=poisson(), offset = lnpop)
summary(ctsfatbasef)
summary(ctsfatf)
plot_model(ctsfatbasef)
plot_model(ctsfatf)
####male
ctsfatbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
ctsfatm <- glm(tot_count ~ age_grp + period+Zvarfat, data=male, family=poisson(), offset = lnpop)
summary(ctsfatbasem)
summary(ctsfatm)
plot_model(ctsfatbasem)
plot_model(ctsfatm)
```
female: `-0.016701` p-value not significant
male: `-0.044007` p-value < 0


## saturated fat-continuous & normalized
```{r, echo=FALSE, message=FALSE}
ctssatfatbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
ctssatfatf <- glm(tot_count ~ age_grp + period+ Zvarsatfat, data=female, family=poisson(), offset = lnpop)
summary(ctssatfatbasef)
summary(ctssatfatf)
plot_model(ctssatfatbasef)
plot_model(ctssatfatf)
####male
ctssatfatbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
ctssatfatm <- glm(tot_count ~ age_grp + period+Zvarsatfat, data=male, family=poisson(), offset = lnpop)
summary(ctssatfatbasem)
summary(ctssatfatm)
plot_model(ctssatfatbasem)
plot_model(ctssatfatm)
```
female: `-0.03444` p-value = 0.1
male: `-0.050498` p-value < 0

## cholesterol-tertiles
```{r cholesterol, echo=FALSE, message=FALSE}
#####female
cholbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
chol1f <- glm(tot_count ~ age_grp + period+chol1mean, data=female, family=poisson(), offset = lnpop)
chol2f <- glm(tot_count ~ age_grp + period+chol2mean, data=female, family=poisson(), offset = lnpop)
chol12f <- glm(tot_count ~ age_grp + period+chol1mean+chol2mean, data=female, family=poisson(), offset = lnpop)
summary(cholbasef)
summary(chol1f)
summary(chol2f)
summary(chol12f)
plot_model(cholbasef)
plot_model(chol1f)
plot_model(chol2f)
plot_model(chol12f)
####male
cholbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
chol1m <- glm(tot_count ~ age_grp + period+chol1mean, data=male, family=poisson(), offset = lnpop)
chol2m <- glm(tot_count ~ age_grp + period+chol2mean, data=male, family=poisson(), offset = lnpop)
chol12m <- glm(tot_count ~ age_grp + period+chol1mean+chol2mean, data=male, family=poisson(), offset = lnpop)
summary(cholbasem)
summary(chol1m)
summary(chol2m)
summary(chol12m)
plot_model(cholbasem)
plot_model(chol1m)
plot_model(chol2m)
plot_model(chol12m)
```
females: tert 2: `-0.0709`, not significant; 
         tert 3 : `-0.258943` not significant
         both: `-0.094288` & `-0.162715` both not significant
males:   tert 2: `-0.011317`, not significant; 
         tert 3 : `-0.315512` not significant
         both: `-0.050203` not significant & `-0.388569` p-value < 0


## continuos alcohol per week- normalized
```{r}

ctsalcbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
ctsalcf <- glm(tot_count ~ age_grp + period+Zvaralcctsmean, data=female, family=poisson(), offset = lnpop)
summary(ctsalcbasef)
summary(ctsalcf)
plot_model(ctsalcbasef)
plot_model(ctsalcf)
####male
ctsalcbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
ctsalcm <- glm(tot_count ~ age_grp + period+Zvaralcctsmean, data=male, family=poisson(), offset = lnpop)
summary(ctsalcbasem)
summary(ctsalcm)
plot_model(ctsalcbasem)
plot_model(ctsalcm)
```
males beta = `0.214506`, p-value not significant, 
females beta = `-0.6534` pvalue = `0.05`
*period 3 is na for males and age groups is na for females? note says model matrix is rank deficient*


## continuous protein- normalized
```{r cts.prot, echo = FALSE, message = FALSE}
ctsprotbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
ctsprotf <- glm(tot_count ~ age_grp + period+Zvarprotmean, data=female, family=poisson(), offset = lnpop)
summary(ctsprotbasef)
summary(ctsprotf)
plot_model(ctsprotbasef)
plot_model(ctsprotf)
####male
ctsprotbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
ctsprotm <- glm(tot_count ~ age_grp + period+Zvarprotmean, data=male, family=poisson(), offset = lnpop)
summary(ctsprotbasem)
summary(ctsprotm)
plot_model(ctsprotbasem)
plot_model(ctsprotm)
```
females: `(-0.044528)` pvalue = 0.05 
males: `(-0.047513)` pvalue < 0.000


## start adding diff RF into model
```{r}
##female
fitf <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
fbmidia <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean, data=female, family=poisson(), offset = lnpop)
fbmidiahbp <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean+hbpmean, data=female, family=poisson(), offset = lnpop)
fthy <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean+hbpmean+thymean1, data=female, family=poisson(), offset = lnpop)
fghp <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean+hbpmean+thymean1+ Zvarghp, data=female, family=poisson(), offset = lnpop)
falc1 <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean+hbpmean+thymean1+ Zvarghp+alcmean1t, data=female, family=poisson(), offset = lnpop)
falc12 <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean+hbpmean+thymean1+ Zvarghp+alcmean1t+alcmean2t, data=female, family=poisson(), offset = lnpop)
fmeatz <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean+thymean1+Zvarghp+Zvarmeatmean, data=female, family=poisson(), offset = lnpop) ## had to remove alc 
summary(fitf)
summary(fbmidia)
summary(fbmidiahbp)
summary(fthy)
summary(fghp)
summary(falc1)
summary(falc12)
summary(fmeatz)

##male
fitm <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
mbmidia <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean, data=male, family=poisson(), offset = lnpop)
mbmidiahbp <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean, data=male, family=poisson(), offset = lnpop)
mthy <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean+thymean1, data=male, family=poisson(), offset = lnpop)
mghp <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean+thymean1+Zvarghp, data=male, family=poisson(), offset = lnpop)
malc1 <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean+thymean1+Zvarghp+alcmean1t, data=male, family=poisson(), offset = lnpop)
malc12 <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean+thymean1+Zvarghp+alcmean1t+alcmean2t, data=male, family=poisson(), offset = lnpop)
mmeatz <- glm(tot_count ~ age_grp + period + bmimean1 + bmimean2 +diabetesmean +hbpmean+thymean1+Zvarghp+alcmean1t+alcmean2t+Zvarmeatmean, data=male, family=poisson(), offset = lnpop) ## this caused NAs for all RF except overweight, gave message: Coefficients: (9 not defined because of singularities)-highly corrrelated variables 


summary(fitm)
summary(mbmidia)
summary(mbmidiahbp)
summary(mthy)
summary(mghp)
summary(malc1)
summary(malc12)
summary(mmeatz)

plot_model(fbmidia)
plot_model(mbmidia)

plot_model(fbmidiahbp)
plot_model(mbmidiahbp)

plot_model(fthy)
plot_model(mthy)

plot_model(mghp)
plot_model(fghp)

plot_model(falc1)
plot_model(falc12)
plot_model(malc1)
plot_model(malc12)

plot_model(fmeatz)
plot_model(mmeatz)

```

**notes:
-female: 
        *diabetes becomes negative once you add in other RF despite being positive in its own model
        *model falc1: age_grp is not defined (NA) due to singularities
        *model falc12: age_grp & alc tert 2 NA due to singularities
        *fmeatz: could not run with alc means in the model, period betas disappear???
-male: 
        *mmeatz model gives message: Coefficients: (9 not defined because of singularities)-obese, diabetes, hbp, thyroid, ghp, alc1&2, meat are          highly correlated apparently