---
title: "9-23 meeting"
author: "lulu zhang"
date: "9/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sas7bdat)
library(tidyr)
library(tidyverse)
library(SASxport)
library(dplyr)
require(haven)
library(car)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
require(ggplot2)
library(janitor)
library(ggplot2)
require(sandwich)
require(msm)
library(forcats)
require(ggplot2)
require(gridExtra)
library(rlang)
library(dplyr)
library(BBmisc)
library(caret)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(ggplot2)
library(readxl)
```


```{r load.data, echo=FALSE, message=FALSE, include=FALSE}
seer = read.csv("./groupedseer.csv") %>% 
  janitor::clean_names()
seerrf=read.csv("./seerrf.csv") %>% 
  mutate(period = as.factor(period))
seergender = read.csv("./seergender.csv")

allrf = read.csv("./seerrf.csv")%>% 
  janitor::clean_names() 

malerf = read.csv("./seermrf.csv")%>% 
  janitor::clean_names()

femrf = read.csv("./seerfrf.csv")%>% 
  janitor::clean_names()

```


```{r seer.inc, echo=FALSE, message=FALSE}
seer %>%
  filter(gender == "Female") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = race)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
    title = "CRC inc rates from 1975-2014 among Females",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
  labels=c(`1` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))

seer %>%mutate(period = as.numeric(period)) %>%
  filter(gender == "Male") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = race)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
    title = "CRC inc rates from 1975-2014 among males ",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(3,4,5,6,7,8,9,10,11),
  labels=c(`3` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))
```

## seer incidence trends among males and females by age groups
```{r, include=FALSE}
seerf = seergender %>% filter(gender == "Female") %>% 
  mutate(period = as.factor(period))
seerm = seergender %>% filter(gender == "Male")%>% 
  mutate(period = as.factor(period))

seerf%>%mutate(period = as.numeric(period),
               age_grp = as.factor(age_grp),
               age_grp = dplyr::recode(age_grp,
                                       `6` = "25-29",
                                       `7` = "30-34",
                                       `8` = "35-39",
                                       `9` = "40-44",
                                       `10` = "45-49"))  %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = age_grp)) +
   labs(
    title = "CRC inc rates from 1975-2014 among females ",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
  labels=c(`1` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))

seerm%>%mutate(period = as.numeric(period),
               age_grp = as.factor(age_grp),
               age_grp = dplyr::recode(age_grp,
                                       `6` = "25-29",
                                       `7` = "30-34",
                                       `8` = "35-39",
                                       `9` = "40-44",
                                       `10` = "45-49"))  %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = age_grp)) +
   labs(
    title = "CRC inc rates from 1975-2014 among males ",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+  scale_x_continuous(breaks=c(1,2,3,4,5,6,7,8,9),
  labels=c(`1` = "1975-1979",`2` = "1980-1984", `3` = "1985-1989", `4`= "1990-1994",
           `5` = "1995-1999", `6` = "2000-2004", `7` = "2005-2009", `8` = "2010-2014", `9` = "2015-2018"))
```

normalize all risk factors for glm!
then plot beta estimates acorss periods, with seer incidence rates overlayed
then another panel with age groups as x-axis 

```{r bmiplot, echo=FALSE, message=FALSE}
## this is just betas from bmi from excel
femalebmi = read_xlsx("./univariate models.xlsx", sheet = 8) %>%
  janitor::clean_names() %>% 
  mutate(beta_estimate = as.numeric(beta_estimate),
         period = as.factor(period))

 age_grp <- femalebmi$age_grp
ggplot(femalebmi, aes(x = period, y = beta_estimate)) + 
  geom_point()+
  labs(title = "beta estimates for period and bmi levels-overweight and obese
       for females")+
geom_hline(yintercept=age_grp, col = "red")

malebmi = read_xlsx("./univariate models.xlsx", sheet = 9) %>%
  janitor::clean_names() %>% 
  mutate(beta_estimate = as.numeric(beta_estimate),
         period = as.factor(period))
 age_grp <- malebmi$age_grp
ggplot(malebmi, aes(x = period, y = beta_estimate)) + 
  geom_point()+
  geom_line(aes(x = period, y = beta_estimate ))+
  labs(title = "beta estimates for period and bmi levels-overweight and obese
       for males") +
geom_hline(yintercept=age_grp, col = "red")

```

## try to overlay seer incidence over beta estimates
This is a plot of the model of beta coefficients for ghp female and male respectively
first and third plot are without rf, second and fourth are with RF
```{r ghp, echo=FALSE, message=FALSE}
ghpctsmean = read.csv("./ghpctsmean.csv")
ghpseer = merge(ghpctsmean, seer, by= c("period", "age_grp", "race", "gender"), all = TRUE) %>% 
  mutate(period = as.factor(period))
ghpseer$period <- relevel(ghpseer$period, ref = "9")
fghp = ghpseer %>% 
  filter(gender == "Female")
mghp = ghpseer %>% 
  filter(gender == "Male")

#female
femghp <- glm(tot_count ~ age_grp + period, data=fghp, family=poisson(), offset = lnpop)
ghpf <- glm(tot_count ~ age_grp + period+ zVar, data=fghp, family=poisson(), offset = lnpop)
summary(femghp)
summary(ghpf)

##plot betas from the model fit
plot_model(femghp, transform = NULL)
plot_model(ghpf, transform = NULL)

#male
maleghp <- glm(tot_count ~ age_grp + period, data=mghp, family=poisson(), offset = lnpop)
ghpm <- glm(tot_count ~ age_grp + period+ zVar, data=mghp, family=poisson(), offset = lnpop)
summary(maleghp)
summary(ghpm)

##plot betas from the model fit
plot_model(maleghp, transform = NULL)
plot_model(ghpm, transform = NULL)



```

plot bmi using plot_model function
```{r bmi.plot}
female = seerrf %>% 
  filter(gender == "Female")
male = seerrf %>% 
  filter(gender == "Male")

female$period <- relevel(female$period, ref = "9")

male$period <- relevel(male$period, ref = "9")
#female
seerbase <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
seerbmi1 <- glm(tot_count ~ age_grp + period+bmimean1, data=female, family=poisson(), offset = lnpop)
seerbmi2 <- glm(tot_count ~ age_grp + period+bmimean2, data=female, family=poisson(), offset = lnpop)
seerbmi23 <- glm(tot_count ~ age_grp + period+bmimean1 +bmimean2, data=female, family=poisson(), offset = lnpop)
summary(seerbase)
summary(seerbmi1)
summary(seerbmi2)
summary(seerbmi23)
plot_model(seerbase)
plot_model(seerbmi1)
plot_model(seerbmi2)
plot_model(seerbmi23)


### males
seerbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
seerbmi1m <- glm(tot_count ~ age_grp + period+bmimean1, data=male, family=poisson(), offset = lnpop)
seerbmi2m <- glm(tot_count ~ age_grp + period+bmimean2, data=male, family=poisson(), offset = lnpop)
seerbmi23m <- glm(tot_count ~ age_grp + period+bmimean1 +bmimean2, data=male, family=poisson(), offset = lnpop)
summary(seerbasem)
summary(seerbmi1m)
summary(seerbmi2m)
summary(seerbmi23m)
##baseline model
plot_model(seerbasem)
##just add overweight
plot_model(seerbmi1m)
## just add obese
plot_model(seerbmi2m)
## add both overweight and obese
plot_model(seerbmi23m)
```



##trying to get two y axis
```{r}
library(plotly)
femalebmi = femalebmi %>% 
  mutate(period = as.factor(period)) 
seerf

x = full_join(femalebmi, seerf) %>% 
  mutate(age_grp = as.factor(age_grp))
obese <- x$obese
 plot_ly(x)%>%
    add_trace(x=~period,y=~rate,name = 'incidence rate',type = 'scatter',mode = 'lines+markers', connectgaps = TRUE, group_by = age_grp) %>%
    add_trace(x=~period,y=~beta_estimate,name = 'beta estimates',type = 'scatter',mode = 'lines+markers',connectgaps = TRUE,yaxis = "y2") %>%
    layout(title = 'bmi beta estimates with incidence rates',
       xaxis = list(title = "incidence rates"),
       yaxis2 = list(side = 'right', overlaying = "y", title = 'beta estimates', showgrid = FALSE, zeroline = FALSE))
 

```

```{r diabetes, echo=FALSE, message=FALSE}
#####female
diabetesbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
diabetesyesf <- glm(tot_count ~ age_grp + period+diabetesmean, data=female, family=poisson(), offset = lnpop)
summary(diabetesbasef)
summary(diabetesyesf)
plot_model(diabetesbasef)
plot_model(diabetesyesf)
####male
diabetesbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
diabetesyesm <- glm(tot_count ~ age_grp + period+diabetesmean, data=male, family=poisson(), offset = lnpop)
summary(diabetesbasem)
summary(diabetesyesm)
plot_model(diabetesbasem)
plot_model(diabetesyesm)
```

## hbp 
```{r hbp}
#####female
hbpbasef <- glm(tot_count ~ age_grp + period, data=female, family=poisson(), offset = lnpop)
hbpyesf <- glm(tot_count ~ age_grp + period+hbpmean, data=female, family=poisson(), offset = lnpop)
summary(hbpbasef)
summary(hbpyesf)
plot_model(hbpbasef)
plot_model(hbpyesf)
####male
hbpbasem <- glm(tot_count ~ age_grp + period, data=male, family=poisson(), offset = lnpop)
hbpyesm <- glm(tot_count ~ age_grp + period+hbpmean, data=male, family=poisson(), offset = lnpop)
summary(hbpbasem)
summary(hbpyesm)
plot_model(hbpbasem)
plot_model(hbpyesm)
```

