---
title: "EDA"
author: "lulu zhang"
date: "7/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sas7bdat)
library(tidyr)
library(tidyverse)
library(SASxport)
library(dplyr)
require(haven)
library(car)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
require(ggplot2)
require(sandwich)
require(msm)
```


```{r incidence.by.race, echo=FALSE,  message=FALSE}
seer = readxl::read_xlsx("./seer.xlsx", sheet = 2) %>% 
  mutate(
    race = as.factor(race),
    gender = as.factor(gender), 
    race = dplyr::recode(race, 
                  `1` = "Black",
                  `0`  = "White",
                  `2` = "Other"),
    gender = dplyr::recode(gender, 
                    `0` = "Male",
                    `1` = "Female"),
    age_grp = as.factor(age_grp),
    age_grp = dplyr::recode(age_grp,
                            `0` = "25-29 year olds",
                            `1` = "30-34 year olds",
                            `2` = "35-39 year olds",
                            `3` = "40-44 year olds",
                            `4` = "45-49 year olds")
  #  period = as.factor(period),
 #  period = dplyr::recode(period,
                #           `0` = "1975-1979",
               #          `1` = "1980-1984",
                 #         `2` = "1985-1989",
                  #        `3` = "1990-1994",
                   #       `4` = "1995-1999",
                    #      `5` = "2000-2004",
                     #    `6` = "2005-2009",
                      #     `7` = "2010-2014")
  )

black_inc = seer %>% 
  filter(race == "Black") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = gender)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
    title = "CRC inc rates from 1975-2014 among Blacks",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

white_inc = seer %>% 
  filter(race == "White") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = gender)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
 title = "CRC inc rates from 1975-2014 among Whites",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

other_inc = seer %>% 
  filter(race == "Other") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = gender)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
 title = "CRC inc rates from 1975-2014 among Other",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  

### add scales = free_y
## add in others
## compare race/ethnicity groups-look at these over time by gender


black_inc
white_inc
other_inc

```


```{r inc.by.gender, echo=FALSE, message=FALSE}
female_inc = seer %>% 
  filter(gender == "Female") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = race)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
    title = "CRC inc rates from 1975-2014 among Females",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

male_inc = seer %>% 
  filter(gender == "Male") %>% 
  ggplot( aes(x = period, y = rate))+
   geom_line(aes(color = race)) +
  facet_wrap(~age_grp, scales = "free_y")+
   labs(
 title = "CRC inc rates from 1975-2014 among males",
    x = "Periods (5-year groups)",
    y = "Inc Rate of CRC",
    caption = "by 5-year age groups (25-49)")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

female_inc
male_inc
```

```{r  echo=FALSE, include=FALSE,  message=FALSE}
mean_wgtsx = read.csv("./rf_wghts.csv")
mean_wgts = mean_wgtsx %>% 
  mutate(
    gender = as.factor(gender),
    race = as.factor(race),
    race = dplyr::recode(race, 
                  `1` = "Black",
                  `0`  = "White",
                  `2` = "Other"),
    gender = dplyr::recode(gender, 
                    `0` = "Male",
                    `1` = "Female"),
     age_grp = as.factor(age_grp),
    age_grp = dplyr::recode(age_grp,
                            `0` = "25-29 year olds",
                            `1` = "30-34 year olds",
                            `2` = "35-39 year olds",
                            `3` = "40-44 year olds",
                            `4` = "45-49 year olds"),
    activity_level..0_Mean = as.numeric(as.character(activity_level..0_Mean)),
    activity_level..1_Mean = as.numeric(as.character(activity_level..1_Mean)),
    activity_level..2_Mean = as.numeric(as.character(activity_level..2_Mean)),
    activity_level..3_Mean = as.numeric(as.character(activity_level..3_Mean)),
     alc_freq..5_Mean= as.numeric(as.character(alc_freq..5_Mean)),
    alc_freq..4_Mean= as.numeric(as.character(alc_freq..4_Mean)),
    alc_freq..3_Mean = as.numeric(as.character(alc_freq..3_Mean)),
    alc_freq..2_Mean = as.numeric(as.character(alc_freq..2_Mean)),
    alc_freq..1_Mean = as.numeric(as.character(alc_freq..1_Mean)),
     alc_freq..0_Mean = as.numeric(as.character(alc_freq..0_Mean)),
    meat_freq_pmo..0_Mean = as.numeric(as.character(meat_freq_pmo..0_Mean)),
    meat_freq_pmo..1_Mean = as.numeric(as.character(meat_freq_pmo..1_Mean)),
    meat_freq_pmo..2_Mean = as.numeric(as.character(meat_freq_pmo..2_Mean)),
    meat_freq_pmo..3_Mean = as.numeric(as.character(meat_freq_pmo..3_Mean)),
    protein..0_Mean = as.numeric(as.character(protein..0_Mean)),
    protein..1_Mean = as.numeric(as.character(protein..1_Mean)),
    protein..2_Mean = as.numeric(as.character(protein..2_Mean)),
    protein..3_Mean = as.numeric(as.character(protein..3_Mean)),
    protein..4_Mean = as.numeric(as.character(protein..4_Mean)),
    fruits_veggies_pmo..0_ = as.numeric(as.character(fruits_veggies_pmo..0_)),
    fruits_veggies_pmo..1_ = as.numeric(as.character(fruits_veggies_pmo..1_)),
    fruits_veggies_pmo..2_ = as.numeric(as.character(fruits_veggies_pmo..2_)),
    fruits_veggies_pmo..3_ = as.numeric(as.character(fruits_veggies_pmo..3_)),
    fruits_veggies_pmo..4_ = as.numeric(as.character(fruits_veggies_pmo..4_)),
    protein..0_Mean = as.numeric(as.character(protein..0_Mean)),
    protein..1_Mean = as.numeric(as.character(protein..1_Mean)),
    protein..2_Mean = as.numeric(as.character(protein..2_Mean)),
    protein..3_Mean = as.numeric(as.character(protein..3_Mean)),
    protein..4_Mean = as.numeric(as.character(protein..4_Mean)),
    sat_fatty_acid.0_Mean = as.numeric(as.character(sat_fatty_acid.0_Mean)),
    sat_fatty_acid.1_Mean = as.numeric(as.character(sat_fatty_acid.1_Mean)),
    sat_fatty_acid.2_Mean = as.numeric(as.character(sat_fatty_acid.2_Mean)),
    sat_fatty_acid.3_Mean = as.numeric(as.character(sat_fatty_acid.3_Mean)),
    sat_fatty_acid.4_Mean = as.numeric(as.character(sat_fatty_acid.4_Mean)),
    still_thyroid_d.0_Mean = as.numeric(as.character(still_thyroid_d.0_Mean)),
    still_thyroid_d.1_Mean = as.numeric(as.character(still_thyroid_d.1_Mean)),
    still_thyroid_d.2_Mean = as.numeric(as.character(still_thyroid_d.2_Mean)),
    age_grp = as.factor( age_grp)
  ) %>% 
  filter(period <=7)

head(mean_wgts)
```

## activity level

```{r activity.level, echo=FALSE,  message=FALSE}
mean_wgts_activity = mean_wgts %>% 
  select(period, age_grp, race, gender, activity_level..0_Mean, activity_level..1_Mean, activity_level..2_Mean,
         activity_level..3_Mean) %>% 
  drop_na()

act_level_0 = mean_wgts_activity%>% 
  ggplot( aes(x = period, y = activity_level..0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender,  scales = "free_y")+
  labs(
 title = "weighted means for little to no activity level by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")



act_level_1 = mean_wgts_activity%>% 
  ggplot( aes(x = period, y = activity_level..1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for moderate activity level by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")


act_level_2 = mean_wgts_activity%>% 
  ggplot( aes(x = period, y = activity_level..2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for high activity level by race and gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")


act_level_3 = mean_wgts_activity%>% 
  ggplot( aes(x = period, y = activity_level..3_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for high activity level by race and gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")



act_level_0
act_level_1
act_level_2


```

## alcohol frequency

```{r alcohol, echo=FALSE,  message=FALSE}

mean_wgts_alc = mean_wgts %>% 
  select(period, age_grp, race, gender, alc_freq..5_Mean,alc_freq..4_Mean,alc_freq..3_Mean,
         alc_freq..2_Mean,alc_freq..1_Mean,alc_freq..0_Mean) %>% drop_na()




alc_freq_4 = mean_wgts_alc%>% 
  ggplot( aes(x = period, y = alc_freq..4_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for alcohol frequency (4 = 7+ times a week) by race & gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")


alc_freq_3 = mean_wgts_alc%>% 
  ggplot( aes(x = period, y = alc_freq..3_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for alcohol frequency (3= 2-3x per week) by race and gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")


alc_freq_2 = mean_wgts_alc%>% 
  ggplot( aes(x = period, y = alc_freq..2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for alcohol frequency (2 = less than once or once a week) by race and gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")


alc_freq_1 = mean_wgts_alc%>% 
  ggplot( aes(x = period, y = alc_freq..1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for alcohol frequency (1 = 3-12x per yr) by race and gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

alc_freq_0 = mean_wgts_alc%>% 
  ggplot( aes(x = period, y = alc_freq..0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for alcohol frequency (0 = none/less than 2-3x per yr) by race and gender",
    x = "Periods (5-year groups)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

alc_freq_0
alc_freq_1
alc_freq_2
alc_freq_3
alc_freq_4

## means do all add up to 1
## every day has mostly 0 because few people responded yes to drinking everyday
#mean_wgts %>% filter(race == "Black", gender == "Female", age_grp == "45-49 year olds")

```




## current smoker, no/yes

```{r current.smoker, echo=FALSE,  message=FALSE}
mean_wgts_smoker = mean_wgts %>% 
  select(period, age_grp, race, gender, current_smkr.0_Mean, current_smkr.1_Mean, current_smkr.2_Mean, 
         current_smkr.3_Mean, current_smkr.NA_Mean) %>% 
  drop_na()


current_smkr_0 = mean_wgts_smoker%>% 
  ggplot( aes(x = period, y = current_smkr.0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for current smoker (0 = no) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

current_smkr_1 = mean_wgts_smoker%>% 
  ggplot( aes(x = period, y = current_smkr.1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for current smoker (1 = yes) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

current_smkr_2 = mean_wgts_smoker%>% 
  ggplot( aes(x = period, y = current_smkr.2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for current smoker (2 = blank but applicable) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

current_smkr_3 = mean_wgts_smoker%>% 
  ggplot( aes(x = period, y = current_smkr.3_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for current smoker (3 = missing) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")



current_smkr_0
current_smkr_1


```

## diabetes y/n

```{r diabetes, echo=FALSE,  message=FALSE}
mean_wgts_diabetes = mean_wgts %>% 
  select(period, age_grp, race, gender,diabetes..0_Mean,diabetes..1_Mean,diabetes..2_Mean ) %>% 
  drop_na()


diabetes_0 = mean_wgts_diabetes%>% 
  ggplot( aes(x = period, y = diabetes..0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for no diabetes by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

diabetes_1 = mean_wgts_diabetes%>% 
  ggplot( aes(x = period, y = diabetes..1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for diabetes = yes by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

diabetes_2 = mean_wgts_diabetes%>% 
  ggplot( aes(x = period, y = diabetes..2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender)+
  labs(
 title = "weighted means for unknown diabetes status by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

diabetes_0
diabetes_1


mean_wgts_diabetes %>% 
  filter(period == 6)
```



## thyroid disease y/n

```{r thyroid.disease, echo=FALSE,  message=FALSE}

thyroid = mean_wgts %>% 
  select(period, age_grp, race, gender, thyroid_d..0_Mean, thyroid_d..1_Mean, thyroid_d..2_Mean) %>% 
  drop_na()

thyroid_0 = thyroid%>% 
  ggplot( aes(x = period, y = thyroid_d..0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for thyroid disease (0 = no) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

thyroid_1 = thyroid%>% 
  ggplot( aes(x = period, y = thyroid_d..1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for thyroid disease (1 = yes) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")



thyroid_0
thyroid_1

```



## home smokers present? 

```{r home.smokers, echo=FALSE, message=FALSE}
home_smoke = mean_wgts%>% 
  select(period, age_grp, race, gender, home_smokers.0_Mean, home_smokers.1_Mean, home_smokers.2_Mean, home_smokers.3_Mean) %>% 
  drop_na()

head(home_smoke)

home_smokers_0 = home_smoke%>% 
  ggplot( aes(x = period, y =  home_smokers.0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for no home smokers by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

home_smokers_1 = home_smoke%>% 
  ggplot( aes(x = period, y =  home_smokers.1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for home smokers = yes by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

home_smokers_0
home_smokers_1
```

## meat consumed per month 

```{r meat_pmo, echo=FALSE, message=FALSE}
meat = mean_wgts %>% 
  select(period, age_grp, race, gender, meat_freq_pmo..0_Mean, meat_freq_pmo..1_Mean,meat_freq_pmo..2_Mean,meat_freq_pmo..3_Mean, meat_freq_pmo..4_Mean, meat_freq_pmo..5_Mean) %>% 
  drop_na() 

meat_0 = meat%>% 
  ggplot( aes(x = period, y =  meat_freq_pmo..0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for meat consumed per month (0 = none) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

meat_1 = meat%>% 
  ggplot( aes(x = period, y =  meat_freq_pmo..1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for meat consumed per month (1 = less than 1x/week) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

meat_2 = meat%>% 
  ggplot( aes(x = period, y =  meat_freq_pmo..2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for meat consumed per month (2 = 1-6x a week) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

meat_3 = meat%>% 
  ggplot( aes(x = period, y =  meat_freq_pmo..3_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for meat consumed per month (3 = 7-154x a week) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

meat_0 
meat_1
meat_2
meat_3
```

## fruits and veggies per month

```{r fruits.veggies, echo=FALSE,  message=FALSE}
fruits_veg = mean_wgts %>% 
  select(period, age_grp, race, gender, fruits_veggies_pmo..0_, fruits_veggies_pmo..1_, fruits_veggies_pmo..2_, fruits_veggies_pmo..3_,fruits_veggies_pmo..4_) %>% 
  drop_na()

fruits_veg_0 = fruits_veg%>% 
  ggplot( aes(x = period, y =  fruits_veggies_pmo..0_) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for fruits & veggies consumed per month (0 = none) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

fruits_veg_1 = fruits_veg%>% 
  ggplot( aes(x = period, y =  fruits_veggies_pmo..1_) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for fruits & veggies consumed per month (1 = less than once a week) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

fruits_veg_2 = fruits_veg%>% 
  ggplot( aes(x = period, y =  fruits_veggies_pmo..2_) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for fruits & veggies consumed per month (2 = 1-6x a week) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

fruits_veg_3 = fruits_veg%>% 
  ggplot( aes(x = period, y =  fruits_veggies_pmo..3_) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for fruits & veggies consumed per month (3 = 7-154x per week) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

fruits_veg_0
fruits_veg_1
fruits_veg_2
fruits_veg_3
```

## protein consumed (categories)

```{r}
protein = mean_wgts %>% 
  select(period, age_grp, race, gender, protein..0_Mean, protein..1_Mean, protein..2_Mean, protein..3_Mean, protein..4_Mean) %>% 
  drop_na()

protein_0 = protein%>% 
  ggplot( aes(x = period, y =  protein..0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of protein consumed (0 = none) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

protein_1 = protein%>% 
  ggplot( aes(x = period, y =  protein..1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of protein consumed (1 = 1-4600g) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

protein_2 = protein%>% 
  ggplot( aes(x = period, y =  protein..2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of protein consumed (2 = 4501-6600g) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

protein_3 = protein%>% 
  ggplot( aes(x = period, y =  protein..3_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of protein consumed (3 = 6601-9200g) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

protein_4 = protein%>% 
  ggplot( aes(x = period, y =  protein..4_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of protein consumed (4 = 9201+) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

protein_0
protein_1
protein_2
protein_3
protein_4

head(protein)
## all adds up to 1
```

## saturated fatty acids (categorical)

```{r echo=FALSE, message=FALSE}
sat_fat = mean_wgts %>% 
  select(period, age_grp, race, gender, sat_fatty_acid.0_Mean, sat_fatty_acid.1_Mean, sat_fatty_acid.2_Mean, sat_fatty_acid.3_Mean,sat_fatty_acid.4_Mean ) %>% 
  drop_na() %>% 
  filter(period == 6)


sat_fat_0 = sat_fat%>% 
  ggplot( aes(x = period, y = sat_fatty_acid.0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of saturated fatty acid consumed (0 = 0) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

sat_fat_1 = sat_fat%>% 
  ggplot( aes(x = period, y = sat_fatty_acid.1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of saturated fatty acid consumed (1 = 1-1500g) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

sat_fat_2= sat_fat%>% 
  ggplot( aes(x = period, y = sat_fatty_acid.2_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of saturated fatty acid consumed (2 = 1501-2400) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

sat_fat_3 = sat_fat%>% 
  ggplot( aes(x = period, y = sat_fatty_acid.3_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of saturated fatty acid consumed (3 =  2401-3600g) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")


sat_fat_4 = sat_fat%>% 
  ggplot( aes(x = period, y = sat_fatty_acid.4_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for grams of saturated fatty acid consumed (4 = 3601+) by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

sat_fat_0
sat_fat_1
sat_fat_2
sat_fat_3
sat_fat_4

head(sat_fat)

## continous nhanes is period 5-7 so it shows up for sat_fat =1 because the ranges were much smaller than the ranges for teh old nhanes
```

## still have thyroid disease y/n

```{r still.thyroid.disease, echo=FALSE, message=FALSE}
still_thyroid = mean_wgts %>% 
  select(period, age_grp, race, gender, still_thyroid_d.0_Mean, still_thyroid_d.1_Mean ) %>% 
  drop_na()


still_thyroid_0 = still_thyroid%>% 
  ggplot( aes(x = period, y = still_thyroid_d.0_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for still thyroid disease = no by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

still_thyroid_1 = still_thyroid%>% 
  ggplot( aes(x = period, y = still_thyroid_d.1_Mean) )+
   geom_line(aes(color = age_grp )) + 
  facet_wrap(~race+gender, scales = "free_y")+
  labs(
 title = "weighted means for still thyroid disease = yes by race & gender",
    x = "Periods (5-year groups from 1975-2014)",
    y = "weighted mean",
    caption = "by 5-year age groups (25-49)")

still_thyroid_0 
still_thyroid_1
```


## continuous variables

```{r load.continuous.data, echo=FALSE,  message=FALSE, include=FALSE}
diet = read.csv("./cts_var_eda.csv") %>% 
  select(-X, -VAR1, -poly_unsat_fa_g)


diet = diet %>% 
 mutate(
    race = as.factor(race),
    gender = as.factor(gender), 
    race = dplyr::recode(race, 
                  `1` = "Black",
                  `0`  = "White", 
                  `2` = "Other"),
    gender = dplyr::recode(gender, 
                    `0` = "Male",
                    `1` = "Female"),
    AGE_GRP = as.factor(AGE_GRP),
    AGE_GRP = dplyr::recode(AGE_GRP,
                            `0` = "25-29 year olds",
                            `1` = "30-34 year olds",
                            `2` = "35-39 year olds",
                            `3` = "40-44 year olds",
                            `4` = "45-49 year olds")
  )

```






## updated continuous variables

```{r echo=FALSE, message=FALSE}
cont = read.csv("./merged2_eda.csv") %>% mutate(
    race = as.factor(race),
    gender = as.factor(gender), 
    race = dplyr::recode(race, 
                  `1` = "Black",
                  `0`  = "White", 
                  `2` = "Other"),
    gender = dplyr::recode(gender, 
                    `0` = "Male",
                    `1` = "Female"),
    AGE_GRP = as.factor(AGE_GRP),
    AGE_GRP = dplyr::recode(AGE_GRP,
                            `0` = "25-29 year olds",
                            `1` = "30-34 year olds",
                            `2` = "35-39 year olds",
                            `3` = "40-44 year olds",
                            `4` = "45-49 year olds"),
    protein_gm = as.numeric(protein_gm),
    sat_fatty_acid = as.numeric(sat_fatty_acid),
    fat_gm = as.numeric(fat_gm),
    fat = coalesce(fat_g, fat_gm),
    protein = coalesce(protein_g, protein_gm),
    saturated_fat = coalesce(saturated_fat_g, sat_fatty_acid),
    genderx = as.factor(genderx),
    racex = as.factor(racex),
    weight1 = as.numeric(weight1),
    strata1 = as.integer(strata1),
    psu1 = as.integer(psu1),
    genderz = coalesce(gender, genderx),
    racez = coalesce(race, racex),
    weight = coalesce(MEC10YR, weight1),
    strata = coalesce(sddsrvyr, strata1),
    cluster = coalesce(sdmvpsu, psu1)
  ) %>% 
select(-protein_g, -protein_gm, -saturated_fat_g, -sat_fatty_acid, -fat_g, -fat_gm, -race, -racex, -gender, -genderx,
       -MEC10YR, -weight1,-sddsrvyr, -strata1,-sdmvpsu, -psu1, -VAR1, -seqn, -corresp_bc,- a_p)

cont = cont %>% 
  rename(gender = genderz,
         race = racez) %>% filter(period <=7)

```







## average drinks per day in the past year

```{r drinks.per.day, echo=FALSE, message=FALSE, include=FALSE}
drinks = cont %>% 
  select(period, AGE_GRP, race, gender, avg_drnks_per_day_pastyr) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period))

drink_p  = ggplot(drinks, aes(x=period, y=avg_drnks_per_day_pastyr,group = AGE_GRP,  color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average drinks consumed per day in the past year from 1975-2014",
    x = "Periods (5-year groups)",
    y = "drinks consumed per day in the past year",
    caption = "by 5-year age groups (25-49)"
  )
drink_p
```

## sugar

```{r sugar , echo=FALSE, message=FALSE, include=FALSE}
sugar = cont %>% 
  select(AGE_GRP, period, race, gender, sugar_gm) %>% 
  mutate(period = as.factor(period)) %>% 
  drop_na() 

sugar_p = ggplot(sugar, aes(x=period, y=sugar_gm,group = AGE_GRP,  color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average sugar in grams consumed from 1975-2014",
    x = "Periods (5-year groups)",
    y = "sugar(g)",
    caption = "by 5-year age groups (25-49)"
  )

sugar_p
```

## fiber 

```{r fiber, echo=FALSE, message=FALSE}
fiber_gm = cont %>% 
  select(period, AGE_GRP, race, gender, fiber_gm) %>% 
  drop_na() %>%
  mutate(period = as.factor(period))

fiber_p = ggplot(fiber_gm, aes(x=period, y=fiber_gm, group = AGE_GRP, color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average fiber intake in grams from 1975-2014",
    x = "Periods (5-year groups)",
    y = "fiber(g)",
    caption = "by 5-year age groups (25-49)"
  )
fiber_p
```
## cholesterol (mg)

```{r cholesterol, echo=FALSE, message=FALSE}
cholesterol = cont %>% 
  select(period, race, gender, AGE_GRP, cholesterol_mg) %>% 
  drop_na() %>% 
  filter(
    race == c("Black", "White")
  )

#diet[, lapply(.SD, mean), by = .(period)]
                
chol_p =  
  ggplot(cholesterol, aes(x=period, y=cholesterol_mg, group = AGE_GRP, color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average cholesterol intake in mg from 1975-2014",
    x = "Periods (5-year groups)",
    y = "cholesterol (mg)",
    caption = "by 5-year age groups (25-49)"
  )
chol_p
```

## sat fat

```{r}
saturated_fat = cont%>% 
  mutate(
    period = as.factor(period)
  ) %>% 
  select(period, race, gender, AGE_GRP, saturated_fat) %>% 
  drop_na() %>%
  filter(race == c("Black", "White"))

saturated_fat_p = ggplot(saturated_fat, aes(x=period, y=saturated_fat, group = AGE_GRP, color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average grams of saturated fat consumed from 1975-2014",
    x = "Periods (5-year groups)",
    y = "saturated fat (g)",
    caption = "by 5-year age groups (25-49)"
  )
saturated_fat_p
```



## mono sat fat

```{r mono.sat.fat, echo=FALSE, message=FALSE}
mono = cont%>% 
  select(period, AGE_GRP, race, gender, mono_sat_fa_g) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period)) %>% 
  filter(race == c("Black", "White"))

mono_p = ggplot(mono, aes(x=period, y=mono_sat_fa_g, group = AGE_GRP, color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average grams of  mono_sat_fa_g consumed from 1975-2014",
    x = "Periods (5-year groups)",
    y = "monosaturated fatty acid (g)",
    caption = "by 5-year age groups (25-49)"
  )

mono_p
```

## old and new nhanes combined-continuous nutrient variables

```{r echo=FALSE, message=FALSE}
cts_var = read.csv("./merge_diet.csv") %>% 
   mutate(
    race = as.factor(race),
    gender = as.factor(gender), 
    race = dplyr::recode(race, 
                  `1` = "Black",
                  `0`  = "White", 
                  `2` = "Other"),
    gender = dplyr::recode(gender, 
                    `0` = "Male",
                    `1` = "Female"),
    AGE_GRP = as.factor(AGE_GRP),
    AGE_GRP = dplyr::recode(AGE_GRP,
                            `0` = "25-29 year olds",
                            `1` = "30-34 year olds",
                            `2` = "35-39 year olds",
                            `3` = "40-44 year olds",
                            `4` = "45-49 year olds"))

```

```{r echo=FALSE, message=FALSE}
protein_cat = cts_var%>% 
  select(period, AGE_GRP, race, gender, protein_g) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period)) 

cat_protein_p = ggplot(protein_cat, aes(x=period, y = protein_g, group = AGE_GRP, color = AGE_GRP)) + 
  stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender, scales = "free_y")+
labs(
    title = "protein intake from 1975-2014",
    x = "period",
    y = "protein (g)",
    caption = "by 5-year age groups (25-49)"
  )
cat_protein_p

```

```{r protein, echo=FALSE, message=FALSE}
prot = cont%>% 
  select(period, AGE_GRP, race, gender, protein) %>% 
  mutate(period = as.factor(period)) %>% 
  drop_na() 

protein_p = ggplot(prot, aes(x=period, y=protein, group = AGE_GRP, color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average protein in grams consumed from 1975-2014",
    x = "Periods (5-year groups)",
    y = "grams of protein",
    caption = "by 5-year age groups (25-49)"
  )
protein_p
```

```{r fat, echo=FALSE, message=FALSE}
fat_cat = cts_var%>% 
  select(period, AGE_GRP, race, gender, fat_g) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period)) 

cat_fat_p = ggplot(fat_cat, aes(x=period, y = fat_g, group = AGE_GRP, color = AGE_GRP)) + 
  stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender, scales = "free_y")+
labs(
    title = "total fat intake from 1975-2014",
    x = "period",
    y = "fat (g)",
    caption = "by 5-year age groups (25-49)"
  )
cat_fat_p
```

## fat

```{r echo=FALSE, message=FALSE}
fat_g = cont %>% 
  select(period, AGE_GRP, race, gender, fat) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period)) %>% 
  filter(race == c("Black", "White"))
  
fat_p = ggplot(fat_g, aes(x=period, y=fat, group = AGE_GRP, color = AGE_GRP)) + stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender)+
labs(
    title = "average grams of  fat consumed from 1975-2014",
    x = "Periods (5-year groups)",
    y = "fat (g)",
    caption = "by 5-year age groups (25-49)"
  )
fat_p
```

```{r echo=FALSE, message=FALSE}
sat_fat_cat = cts_var%>% 
  select(period, AGE_GRP, race, gender, saturated_fat_g) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period)) 

sat_fat_p = ggplot(sat_fat_cat, aes(x=period, y = saturated_fat_g, group = AGE_GRP, color = AGE_GRP)) + 
  stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender, scales = "free_y")+
labs(
    title = "total saturated fat intake from 1975-2014",
    x = "period",
    y = "saturated fat (g)",
    caption = "by 5-year age groups (25-49)"
  )
sat_fat_p

sat_fat_cat %>% 
  filter(period == 2)
```

```{r echo=FALSE, message=FALSE}
cholest_cat = cts_var%>% 
  select(period, AGE_GRP, race, gender, cholesterol_mg) %>% 
  drop_na() %>% 
  mutate(period = as.factor(period)) 

cholest_p = ggplot(cholest_cat, aes(x=period, y = cholesterol_mg, group = AGE_GRP, color = AGE_GRP)) + 
  stat_summary(fun.y="mean", geom="line")+
 facet_wrap(~race+gender, scales = "free_y")+
labs(
    title = "total cholesterol intake from 1975-2014",
    x = "period",
    y = "cholesterol_mg",
    caption = "by 5-year age groups (25-49)"
  )
cholest_p
```

