---
title: "pre merge plots"
author: "lulu zhang"
date: "8/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sas7bdat)
library(tidyr)
library(tidyverse)
library(SASxport)
library(dplyr)
require(haven)
library(car)
library(readxl)
library(here)
library(skimr) # install.packages('skimr')
library(kableExtra) # install.packages('kableExtra')
require(ggplot2)
library(janitor)
library(ggplot2)
require(sandwich)
require(msm)
library(forcats)
require(ggplot2)
require(gridExtra)
library(naniar)
```


```{r load.data, include=FALSE}
new_nhanes = read.csv("./new_nhanes_diet_per.csv") %>% 
  select(-sat_fatty_acid, -cholesterol, -protein, -fat, -fiber, -sugar, -wtdrd1, -wtdr2d)
new_diet = read.csv("./new_nhanes_tot_diet.csv") %>% 
  select(seqn, protein, fiber, fat, sat_fatty_acid, cholesterol, sugar, wtdrd1, wtdr2d)
h1ac = read.csv("./h1ac.csv") %>% 
  select(-X) %>% 
  janitor::clean_names() %>% 
  rename(h1ac = lbxgh)

growth_chrts = read.csv("./growth_chts_63_94.csv")  #1963-1994

new_nhanes = merge(new_nhanes, new_diet, by = "seqn")
new_nhanes = merge(new_nhanes, h1ac, by = "seqn")

nhanesI = read.csv("./nhanesI_per.csv")
nhanesII = read.csv("./nhanesII_per.csv")
nhanesIII = read.csv("./nhanesIII_per.csv")
HEMO_N3 = read.csv("./N3_HEMO.CSV") %>% 
  janitor::clean_names()

nhanesIII = merge(nhanesIII, HEMO_N3) 
nhanesI_diet = read.csv("./nhanesI_diet_per.csv")  ## has dob, month (01-12), year, (1986-1962) 
nhesI = read.csv("./nhesI_per.csv")
n1_htwt = read.csv("./n1_ht_wt.csv")
n3_diet = read.csv("./n3_diet.csv")
```

```{r tidy.new.nhanes, message=FALSE, echo=FALSE, include=FALSE}
new_nhanes = new_nhanes %>% 
  mutate(gender = as.factor(gender),
         gender = dplyr::recode(gender, 
                                `0` = "Male",
                                `1` = "Female"),
         race = as.factor(race),
         race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
         AGE_GRP = as.factor(AGE_GRP),
         AGE_GRP = dplyr::recode(AGE_GRP,
                                 `0` = "25-29",
                                 `1` = "30-34",
                                 `2` = "35-39",
                                 `3` = "40-44",
                                 `4` = "45-49"))
nhanesI_diet = nhanesI_diet %>% 
  mutate(gender = as.factor(gender),
         gender = dplyr::recode(gender, 
                                `0` = "Male",
                                `1` = "Female"),
         race = as.factor(race),
         race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
         AGE_GRP = as.factor(AGE_GRP),
         AGE_GRP = dplyr::recode(AGE_GRP,
                                 `0` = "25-29",
                                 `1` = "30-34",
                                 `2` = "35-39",
                                 `3` = "40-44",
                              `4` = "45-49"),
        cholesterol_mg = as.numeric(cholesterol_mg)/100,
         tot_sat_fat = as.numeric(tot_sat_fat)/100,
         fat = as.numeric(fat)/100,
         protein = as.numeric(protein)/100
         ) %>% 
  rename( sat_fatty_acid = tot_sat_fat ,
         cholesterol = cholesterol_mg)


nhanesI = nhanesI  %>% 
  mutate(gender = as.factor(gender),
         gender = dplyr::recode(gender, 
                                `0` = "Male",
                                `1` = "Female"),
         race = as.factor(race),
         race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
         AGE_GRP = as.factor(AGE_GRP),
         AGE_GRP = dplyr::recode(AGE_GRP,
                                 `0` = "25-29",
                                 `1` = "30-34",
                                 `2` = "35-39",
                                 `3` = "40-44",
                              `4` = "45-49"))
n1_htwt = n1_htwt  %>% 
  janitor::clean_names()%>% 
  mutate(
    height_cm = as.numeric(levels(height_cm))[height_cm]/10,
    weight_kg = as.numeric(levels(weight_kg))[weight_kg]/100)
n1_htwt$height_m <- n1_htwt$height_cm/100
n1_htwt$bmi <- (n1_htwt$weight_kg/(n1_htwt$height_m^2))

nhanesi = merge(nhanesI, n1_htwt, by= "seqn") 
 nhanesI_diet 
```

```{r message=FALSE, echo=FALSE, include=FALSE}
nhanesII= nhanesII %>% 
  mutate(gender = as.factor(gender),
         gender = dplyr::recode(gender, 
                                `0` = "Male",
                                `1` = "Female"),
         race = as.factor(race),
         race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
         AGE_GRP = as.factor(AGE_GRP),
         AGE_GRP = dplyr::recode(AGE_GRP,
                                 `0` = "25-29",
                                 `1` = "30-34",
                                 `2` = "35-39",
                                 `3` = "40-44",
                                 `4` = "45-49"),
         period = as.numeric(period),
         height_ft =as.numeric(levels(height_ft))[height_ft],   ###888:999 NA
         height_in = as.integer(levels(height_in))[height_in],
         weight_lb = as.integer(levels(weight_lb))[weight_lb])%>% 
  replace_with_na(replace = list(weight_lb = 888)) %>% 
  replace_with_na(replace = list(height_in = 88))%>% 
  replace_with_na(replace = list(height_ft = 8))
 
nhanesII$height_cm <- (nhanesII$height_ft*12+nhanesII$height_in)*2.54
nhanesII$weight_kg <- nhanesII$weight_lb/2.205
## calc bmi frim height and weight
nhanesII$height_m <- nhanesII$height_cm/100
nhanesII$bmi <- (nhanesII$weight_kg/(nhanesII$height_m^2))
nhanesII

nhanesIII = nhanesIII%>% 
  mutate(gender = as.factor(gender),
         gender = dplyr::recode(gender, 
                                `1` = "Male",
                                `2` = "Female"),
         race = as.factor(race),
         race = dplyr::recode(race,
                              `0` = "White",
                              `1` = "Black",
                              `2` = "Other"),
         AGE_GRP = as.factor(AGE_GRP),
         AGE_GRP = dplyr::recode(AGE_GRP,
                                 `0` = "25-29",
                                 `1` = "30-34",
                                 `2` = "35-39",
                                 `3` = "40-44",
                                 `4` = "45-49"),
         period = as.numeric(period)) %>% 
  replace_with_na(replace = list(weight_lb = 999)) %>% #888:999 are NA
  replace_with_na(replace = list(weight_lb = 888))%>% 
  replace_with_na(replace = list(height_in = 888))%>% 
  replace_with_na(replace = list(height_in = 888))

 
nhanesIII$height_cm <- nhanesIII$height_in*2.54
nhanesIII$weight_kg <- nhanesIII$weight_lb/2.205
## calc bmi frim height and weight
nhanesIII$height_m <- nhanesIII$height_cm/100
nhanesIII$bmi <- (nhanesIII$weight_kg/(nhanesIII$height_m^2))
nhanesIII 
n3_diet = n3_diet %>% 
  janitor::clean_names()

nhanesIII = merge(nhanesIII, n3_diet, by = "seqn")
```

## combine new nhanes, nhanes II, nhanesIII

```{r merge, include=FALSE}
new_nhanesx = new_nhanes %>% 
  select(gender, race, period, AGE_GRP, bmi, weight_kg, height_cm )

nhanesix = nhanesi %>% 
  select(gender, race, period, AGE_GRP, bmi, weight_kg, height_cm )

nhanesIIx = nhanesII %>% 
  select(gender, race, period, AGE_GRP, bmi, weight_kg, height_cm )

nhanesIIIx = nhanesIII %>% 
  select(gender, race, period, AGE_GRP, bmi, weight_kg, height_cm )

merged1 = rbind(nhanesix, nhanesIIx)
merged2 = rbind(nhanesIIIx, new_nhanesx)
merged = rbind(merged1, merged2)

```


### plotting BMI 

```{r bmi.plot, message=FALSE, echo=FALSE}
# histogram to look at distribution
new_bmi = merged %>% 
  filter(period == 2)
hist( x = new_bmi$bmi)

#plot avg bmi across periods for new nhanes (1999-2018)
bmi_p2 = merged %>%  
  filter(between (period, 0,7),
         between (bmi ,12,80)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_bmi = mean(bmi, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_bmi, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg BMI from 1975-2014 among 25-49 year olds",
       ylab = "average BMI")

bmi_p2 +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

### no data for period 2 (1985-1989)
```



```{r height, echo=FALSE, message=FALSE}

height_c = merged %>% 
  filter(period %in% c(4:5),
         race == "White",
         gender == "Female",
         between(height_cm, 20, 220))
        # between(height_cm, 20, 270))

hist( x = height_c$height_cm)

#plot avg bmi across periods for new nhanes (1999-2018)
height_p2 = merged %>%  
  filter(between(period,0,7),
         between(height_cm ,20, 250)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_height_cm = mean(height_cm, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_height_cm, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg height (cm) from 1975-2014 among 25-49 year olds",
       ylab = "average height(cm)")

height_p2 +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))



```

Male, Other race, 25-29 year olds (red line) looks odd for period 1 because only 2 observations, heights = 162.56, 157.48


```{r weight, message=FALSE, echo=FALSE}
weight_kg = merged %>% 
  filter(period <= 5,
         gender == "Female",
         race == "Black",
         AGE_GRP == "35-39")

hist( x = weight_kg$weight_kg)

#plot avg bmi across periods for new nhanes (1999-2018)
weight_p2 = merged %>%  
  filter(between(period,0,7),
         between(weight_kg ,10, 250)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_weight_kg = mean(weight_kg, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_weight_kg, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg weight (kg) from 1975-2014 among 25-49 year olds",
       ylab = "average weight (kg)")

weight_p2 +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```

```{r cts.diet.var, message=FALSE, echo=FALSE, include=FALSE}
n1_diet = nhanesI_diet %>% 
  select(seqn, race, gender, AGE_GRP, period, cholesterol, sat_fatty_acid, fat, protein) %>% 
  filter(period <=7)
nhanesi

nhanesIId = nhanesII %>% 
  mutate(
    tot_sat_fat = as.numeric(tot_sat_fat)/100,
    fat = as.numeric(fat)/100,
    protein = as.numeric(protein)/100 ) %>%
  rename(sat_fatty_acid = tot_sat_fat) %>% 
  select(seqn, race, gender, AGE_GRP, period, sat_fatty_acid, fat, protein) %>% 
  filter(period <= 7)

nhanes3d = nhanesIII %>% select(seqn, race, gender, AGE_GRP, period, cholesterol, sat_fatty_acid, fat, protein) 

new_nhanes_diet = new_nhanes %>% select(seqn, race, gender, AGE_GRP, period, cholesterol, sat_fatty_acid, fat, protein, fiber, sugar) %>% distinct() 

d1 = full_join(n1_diet, nhanesIId)
d2= full_join(nhanes3d, new_nhanes_diet)
diet = full_join(d1, d2)

diet = diet %>% 
  replace_with_na(replace = list(sat_fatty_acid = 999999.99))%>% 
  replace_with_na(replace = list(cholesterol = 8888.0))

```

```{r fat, echo=FALSE, message=FALSE}
fat_h =  diet %>% 
  filter(period <= 1,
         gender == "Male",
         race == "Other",
         AGE_GRP == "25-29",
         between(fat, 0,500))

hist( x = fat_h$fat)

#plot avg bmi across periods for new nhanes (1999-2018)
fat_p = diet %>%  
  filter(between(period,0,7),
         between(fat ,0, 500)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_fat = mean(fat, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_fat, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg fat (g) intake  from 1975-2014 among 25-49 year olds",
       ylab = "average fat (g)")

fat_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

Other, Male, 25-29 year olds (red) has 11 observations, 3 observations had very high values, fat = 217.64, 114.72, 113.21-probs why there is a spike

```{r sat.fat, echo=FALSE, message=FALSE}

satfat_h =  diet %>% 
  filter(period <=4,
         AGE_GRP == "35-39",
         race == "White",
         gender == "Male",
         between(sat_fatty_acid, 0,200))

#hist( x = satfat_h$sat_fatty_acid)

#plot avg bmi across periods for new nhanes (1999-2018)
satfat_p = diet %>%  
  filter(between(period,0,7),
         between(sat_fatty_acid ,0, 200)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_satfat = mean(sat_fatty_acid, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_satfat, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg saturated fat (g) intake  from 1975-2014 among 25-49 year olds",
       ylab = "average saturated fat (g)")

satfat_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r protein, echo=FALSE, message=FALSE}
prot_h =  diet %>% 
  filter(period <=4,
         AGE_GRP == "40-44",
         race == "Other",
         gender == "Female",
         between(protein, 0,300))

#hist( x = prot_h$protein)

#plot avg bmi across periods for new nhanes (1999-2018)
prot_p = diet %>%  
  filter(between(period,0,7),
         between(protein ,0, 300)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_prot = mean(protein, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_prot, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg protein (g) intake  from 1975-2014 among 25-49 year olds",
       ylab = "average protein fat (g)")

prot_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r cholesterol.plots, echo=FALSE, message=FALSE}
chol_h =  diet %>% 
  filter(period <=7,
       AGE_GRP == "35-39",
        race == "Other",
       gender == "Male",
         between(cholesterol, 0,3000))

#hist( x = chol_h$cholesterol)

#plot avg bmi across periods for new nhanes (1999-2018)
chol_p = diet %>%  
  filter(between(period,0,7),
         between(cholesterol ,0, 300)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_chol = mean(cholesterol, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_chol, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg cholesterol (mg) intake  from 1975-2014 among 25-49 year olds",
       ylab = "average cholesterol (mg)")

chol_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## only new nhanes (1999-2018) has fiber data

```{r fiber.plot, echo=FALSE, message=FALSE}

fiber_h =  diet %>% 
  filter(period <=7,
      #   AGE_GRP == "40-44",
       #  race == "Other",
        # gender == "Female"
         between(fiber, 0,150))

#hist( x = fiber_h$fiber)

#plot avg bmi across periods for new nhanes (1999-2018)
fiber_p = diet %>%  
  filter(between(period,0,7),
         between(fiber ,0, 150)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_fiber = mean(fiber, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_fiber, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg fiber (g) intake  from 1999-2014 among 25-49 year olds",
       ylab = "average fiber (g)")

fiber_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r sugar.plot, echo=FALSE, message=FALSE}

sugar_h =  diet %>% 
  filter(period ==6,
      #   AGE_GRP == "40-44",
       #  race == "Other",
        # gender == "Female"
         between(sugar, 0,350))

#hist( x = sugar_h$sugar)

#plot avg bmi across periods for new nhanes (1999-2018)
sugar_p = diet %>%  
  filter(between(period,0,7),
         between(sugar ,0, 150)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_sugar = mean(sugar, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_sugar, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg sugar (g) intake  from 1999-2014 among 25-49 year olds",
       ylab = "average sugar (g)")

sugar_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


## glycohemoglobin (%) data available for NHANESIII + continuous NHANES

```{r glyco, echo=FALSE, message=FALSE}

nhanesIII_hemo = nhanesIII %>% 
  select(AGE_GRP, period, race, gender, ghp)

new_nhanes_hemo = new_nhanes %>% 
  select(AGE_GRP, period, race, gender, h1ac) %>% 
  rename(ghp = h1ac)

glyco = rbind(nhanesIII_hemo, new_nhanes_hemo) %>% 
  mutate(
    AGE_GRP = as.factor(AGE_GRP),
    AGE_GRP = dplyr::recode(AGE_GRP, 
                        `0` = "25-29",
                        `1` = "30-34",
                        `2` = "35-39",
                        `3` = "40-44",
                        `4` = "45-49"),
    race = as.factor(race),
    race = dplyr::recode(race,
                         `0` = "White",
                         `1` = "Black",
                         `2` = "Other"),
    gender = as.factor(gender),
    gender = dplyr::recode(gender,
                           `1` = "Male",
                           `2` = "Female"))


glyco_p = glyco %>%  
  filter(between(period,0,7),
         between(ghp ,0, 17)) %>% 
           group_by( period, AGE_GRP, race, gender) %>% 
           summarize(avg_h1ac = mean(ghp, na.rm = TRUE)) %>% 
ggplot(aes(x=period, y=avg_h1ac, color=AGE_GRP))+
  geom_line()+
  facet_wrap(~race+gender, scales = "free_y")+
  labs(title = "Avg glycohemoglobin (%) from 1999-2014 among 25-49 year olds",
       ylab = "average % glygohemoglobin")

glyco_p +  scale_x_continuous(breaks=c(0, 1,2,3,4,5,6,7),
  labels=c( `0` = "1974-1979",
           `1` = "1980-1984",
                             `2` = "1985-1989",`3` = "1990-1994",
                           `4` = "1995-1999", `5` = "2000-2004",
                           `6` = "2005-2009", `7` = "2010-2014"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## growth charts for ages 0-25 year olds from 1963-1994

```{r bmi, echo=FALSE, message=FALSE}
growth_chrts = growth_chrts %>% 
  mutate(
    survey = as.factor(survey))


growth_chrts  <- growth_chrts %>% mutate(survey_yr = case_when(survey == 1 ~ "1963-1965 (NHESII)",
                               survey == 2 ~ "1966-1970 (NHESIII)",
                               survey == 3 ~ "1971-1974 (NHANESI)",
                               survey == 4 ~ "1976-1980 (NHANESII)",
                               survey == 5 ~ "1988-1994(NHANESIII)"),
         age_grp = case_when(age >= 0  & age <= 5 ~ '0-5',
                             age >= 6  & age <= 10 ~ '6-10',
                             age >= 11  & age <= 15 ~ '11-15',
                             age >= 16  & age <= 20 ~ '16-20',
                             age >= 21  & age <= 25 ~ '21-25'),
         age_grp = as.factor(age_grp), 
         survey_yr = as.factor(survey_yr),
         survey = as.numeric(survey))

growth_chrts %>% 
  filter(between(bmi,6,70)) %>% 
           group_by( age_grp, survey, gender) %>% 
           summarize(avg_bmi= mean(bmi, na.rm = TRUE)) %>% 
ggplot(aes(x=survey, y=avg_bmi, color=age_grp))+
  geom_line()+
  facet_wrap(~gender, scales = "free_y")+
  labs(title = "Avg bmi from 1963-1994 among 0-25 year olds",
       ylab = "average bmi") +  scale_x_continuous(breaks=c(1,2,3,4,5),
  labels=c(`1` = "1963-1965(NHESII)",
           `2` = "1966-1970(NHESIII)",
           `3` = "1971-1974(NHANESI)",`4` = "1976-1980(NHANESII)", `5` = "1988-1994(NHANESIII)"))+
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


## can filter by survey year and then merge to corresp nhanes/nhes via seqn number

```
NHESII:6-11 year olds
NHESIII:12-17 year olds 

```{r separate.by.survey, include=FALSE}

nhesII_gc = growth_chrts %>% 
  filter(survey == 1)

nhesIII_gc = growth_chrts %>% 
  filter(survey == 2)

nhanesI_gc = growth_chrts %>% 
  filter(survey == 3) %>% 
  select(-X)
write.csv(nhesIII_gc, "./nhesIII_gc.csv")
nhanesII_gc = growth_chrts %>% 
  filter(survey == 4)

nhanesIII_gc = growth_chrts %>% 
  filter(survey == 5)

nhanesIx = read.csv("./nhanesI.csv") %>% 
  select(-X, age, race, gender, bmi)

nhanesII = read.csv("./nhanesII.csv") %>% 
select(-X)

nhanesIx

n3 = read.csv("./n3.csv") %>% 
  select(-X)

nhesIII = read.csv("./nhesIII.csv") %>% 
  select(-X)
nhesII = read.csv("./nhesII.csv") %>% 
  select(-X)
```


```{r old.growth.data, echo=FALSE, include=FALSE}
nhesII
nhesIII
nhanesIx
nhanesII
n3
```

